### 2020-06-10 Covid_ethnicity_Scotlandhc.Rmd

---
title: "Briefing paper on Ethnicity and COVID-19 status in Scotland"

author: Modelling & Research Cell, Public Health Scotland

date: "`r gsub('^0', '', format(Sys.time(), '%d %B %Y'))`"

header-includes: |
  \let\origquote\quote
  \def\quote{\origquote\itshape}
  \usepackage{graphicx}
  \usepackage{longtable}
  \usepackage{booktabs}
  \usepackage{float}
  \usepackage{array}
  \usepackage{longtable}
  \usepackage{lscape}

output:
  pdf_document:
  keep_tex: true
  latex_engine: lualatex

# rmarkdown::render("Covid_ethnicity_Scotland.Rmd")

---

# Background
There have been reports of an excess of hospitalisations, intensive care unit need and death from COVID-19 among non-white ethnic groups in particular Black or Black British and South Asian.   For example a preliminary study using UK Biobank data has [reported](http://medrxiv.org/cgi/content/short/2020.04.22.20075663) that Black, South Asian and White Irish people were more likely to have confirmed infection (RR 4.01 (95% CI 2.92-5.12); RR 2.11 (95% CI 1.43-3.10); and RR 1.60 (95% CI 1.08-2.38) respectively).

Understanding whether such inequity in risk of COVID-19 by ethnicity pertains in Scotland and whether disease severity is worse is relevant to early anticipation of health care needs, triaging algorithms being used by NHS Hubs, and shielding policies.

The purpose of this report was to  i) enumerate the numbers of cases of  COVID-19 of non-White ethnicity ii) to examine whether there is any evidence of ethnic differences in COVID-19 disease in Scotland acknowledging that the size of non-White ethnic minority populations is small in Scotland and that ethnicity is not known with completeness.

# Methods 
The REACT-SCOT case-control study was set up to investigate risk factors for severe COVID-19 and to lay the basis for developing risk stratification algorithms. All persons to date  who tested positive  for SARS-CoV-2 nucleic acid were identified using the ECOSS virology laboratory systems. This system records CHI but not ethnicity.

Any additional persons with a registered death with a COVID-19 code ( suspected or probable ) were also identified.

A matched set of 10 controls per case matched for age, sex and GP practice was identified using the CHI database as part of the REACT-COVID19 case control study led by the Modelling & Research cell.

Data were extracted for  this analysis on 15th May 2020. They do not include the Lighthouse data as we do not have access to this through ECOSS as yet.

## Coding of ethnicity
Names of cases and controls were available to the CHI team. They used the ONOMAP software to assign predicted ethnicity to all cases and controls. This was output in anonymised format to the Modelling & Research Cell linkage team. ONOMAP is a software package developed by Mateos at UCL which used people's first and family names to estimate underlying ethnic group. It has been developed using algorithms built from the Worldnames database that covered both names and ethnic group in a large number of countries. ONOMAP assigns both predicted geographical region of origin and onolytic type. ONOMAP has previously been reported to have variable sensitivity for assigning non-White ethnic groups. ([Lakha 2011](https://www.ncbi.nlm.nih.gov/pubmed/21907365)) The ONOMAP "Onolytic Types" were used to assign ethnicity. Names classified "Muslim" in the ONOMAP classification were assigned to the South Asian group, as Scotland most people with Muslim names are South Asian.  The ONOMAP classification assigns a high proportion of names in Scotland to the category "Celtic" which does not distinguish Irish from Scottish names so that White Irish cannot be differentiated from White British. A previous comparison of ONOMAP against SMR01 records in Scotland concluded that SMR01 records with a missing ethnic group code were broadly representative of records with ethnicity recorded (Clark & Buchanan ISD 2017).

Hospital discharge and outpatient attendance data (SMR01 and SMR00 respectively) have increased in their completion of  recording of self assigned ethnic group in recent years.   Census categories are used. However there remains substantial missingness.   For all cases and controls the linkage team extracted i) the most recent ethnic group assignation from these two datasets  with a 10 year lookback period. Unknown or missing codes were excluded.  Concordance between the most recent SMR ethnic group   and the most frequent  SMR ethnic group  recorded was 98%.

The classification given below follows the Census categories, collapsed into five groups: White, Black, South Asian, Chinese, Other.  The Black category combines those who self-identify as Black British, Black Caribbean, Black African and Black Other.

The rate ratios shown are based on conditional logistic regression.  With the incidence density sampling used in this study, the odds ratios in a conditional logistic regression are equal to rate ratios.  The unconditional odds ratios calculated from the numbers of cases and controls in the tables are not the same as the conditional odds ratios and should not be used.

As other sociodemographic factors may vary by ethnicity we further adjusted analyses for residential care home status and Scottish Index of Multiple Deprivation (SIMD).

# Results
SMR based ethnicity was unavailable for approximately 25% of cases and 50% of controls. The ONOMAP approach  assigned ethnicity  in >99% of persons.  Given the substantial and differential missingness of SMR based ethnicity the principle value of showing the  SMR data here is to  inform on the likely absolute  numbers of cases   of non-White ethnicity rather than for the purposes of estimating rate ratios.

## Agreement between ethnicity assigned by name classification (ONOMAP) and self-reported ethnicity in  SMR records

Table \ref{tab:crosstab} shows the agreement between the ONOMAP  (name based) assigned ethnicity and the self assigned SMR ethnicity.  The name classification assigns to the same as  SMR category `r table.ethnic[1, 1]` of those who self-report their ethnicity as White, `r table.ethnic[5, 5]` of those who self-report as South Asian, `r table.ethnic[3, 3]` of those who self-report as Chinese, but only `r table.ethnic[2, 2]` of those who self-report as Black.  The proportion with  non-White ethnicity  is low for all groups. Thus  although the sensitivity of ONOMAP for assigning SMR South Asian ethnicity is >99% the positive predictive value of ONOMAP assigned South Asian ethnicity is just 60%.

Clearly analyses that distinguish the Black group can be done only for the subset of individuals who have SMR records.  *Please note that we do not consider the rate ratios here to be reliable because of the differential missingness.* Therefore in all further analyses, the ONOMAP classification has been collapsed to four groups: White, South Asian, Chinese, Other.

## Relation of test -positive, hospitalised and severe / fatal COVID-19 to ethnicity in the Scottish population
The tables below compare cases and controls  first with ethnicity based on SMR classification for

i)all test positive cases ii) all hospitalized or severe  test positive cases, iii) severe  test positive cases iv) severe test positive cases OR a COVID-19 death registration,    Severe test positive cases are those who  entered a critical care unit  or died within 28 days of a positive test.  Note that full follow up to 28 days has not yet completed for  the recent cases. The first column of rate ratios are univariate. Those in the second column are further adjusted for SIMD and care home status.

Subsequent tables show the same comparisons with ethnicity based on ONOMAP name classification. Tables \ref{tab:severity.smr} and \ref{tab:severity} show the distribution over severity categories by ethnicity. Tables \ref{tab:carehome} and \ref{tab:board} show the distribution of care home status and health board.  Note that  p -values  that are so extreme that they are less than the threshold for printing a floating point number are output as 0.

Tables comparing all test-positive individuals with controls and hospitalised test  positives are included by request.   However they should be interpreted with caution, as selection for testing is influenced by many factors that are changing through the epidemic as policy changes.  Similar considerations apply to proportion with severe  disease  among those who test positive.   **We  therefore strongly advise that the most useful data  for examining ethnic differences are those for severe test positive disease  and registered deaths as they will have least selection effects.**

There was no evidence of elevation in risk of  "severe  test positive disease" or "severe test positive disease or death"  in  any non-White ethnic group using  either the ONOMAP  or SMR  approach.

Using the named based classification for which there are much more data  for those of South Asian ethnicity the rate ratio was below 1  and was statistically significant.  A further disaggregation of the name based classification into Muslim South Asian and Non-Muslim South Asian ( data not presented) showed risk ratios of similar size below 1 for both subgroups.  For those of Chinese  ethnicity the name based classification showed lower test positive rates also.

Adjusting for SIMD and residential care home status  did not affect these estimates.

There were increased risk ratios for non-White ethnic groups for test positivity  using SMR  - however note our caveats on interpretation of   this - both because of differential missingness in SMR based ethnicity and because of selection effects on test positive levels.

# Conclusion

With current data obtaining a complete picture of the ethnic distribution of  people with COVID-19 disease in Scotland is difficult.  Historical medical records are incomplete and differentially so in their recording of ethnicity.  The numbers of  people in non-White ethnic groups in our population is relatively low compared with England and not known with certainty. Our attempt to improve completeness of ethnicity assignation is of some use for those of South Asian and Chinese origin but remains imperfect and  is not useful for those of Black ethnicity.

Here for transparency we have  nonetheless presented the results of analyses we specified in advance that we would undertake.  However there are important caveats on interpreting these data as we have outlined above.

The absolute numbers shown by ethnicity are useful in providing some indication of the likely  upper bound on numbers of cases in non-White ethnic groups.

We do not find evidence at this stage of elevations in risk of severe COVID-19 disease in non-White ethnic groups compared to White.

Further approaches need to be explored to improve ascertainment of ethnicity. One possibility may be to train a new  Scotland specific name based classifier against the SMR  based data and then apply that.

\newpage

```{r settings, echo = FALSE, results='asis', warning=FALSE, error=FALSE}

# GLobal R options for this document
options(knitr.table.format="latex")
options(knitr.kable.NA='')

```

# Tables

```{r ethnic, echo=FALSE, warning=FALSE, message=FALSE}

knitr::kable(table.ethnic, 
             escape=FALSE, 
             booktabs=TRUE,
             caption="\\label{tab:crosstab}Comparison of ethnicity assigned by ONOMAP name classification (rows) with ethnicity assigned in SMR  records (columns)") %>%
    add_header_above(c(" "=1, "SMR records"= 6)) %>%
    pack_rows("ONOMAP", 1, 5) %>%
    kable_styling(latex_options="hold_position")

```

```{r ethnicsmr.testpos, echo=FALSE, warning=FALSE, message=FALSE}

rownames(table.testpositives.demog.ethnicsmr) <-
    gsub("_", " ", rownames(table.testpositives.demog.ethnicsmr))
knitr::kable(table.testpositives.demog.ethnicsmr,
             escape=FALSE, 
             booktabs=TRUE,
             label="testpos.ethnicsmr",
             col.names=c(clean.header(colnames(table.testpositives.demog.ethnicsmr)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6), 
             caption="Associations of test-positive status with ethnicity in those with ethnicity recorded in Scottish Morbidity record") %>%
    add_header_above(c(" "=3, "Univariate"=2, "Multivariate"=2)) %>%
    kable_styling(latex_options="hold_position") %>% 
    column_spec(column=1, width="2.7cm")  %>%
    column_spec(column=2, width="2cm")  %>%
    column_spec(column=3, width="1.85cm")  %>%
    column_spec(column=4, width="2.5cm") %>%
    column_spec(column=5, width="1.5cm") %>%
    column_spec(column=6, width="2.5cm") %>%
    column_spec(column=7, width="1.5cm")
```

```{r ethnicsmr.hosp, echo=FALSE, warning=FALSE, message=FALSE}

rownames(table.hospitalized.demog.ethnicsmr) <-
    gsub("_", " ", rownames(table.hospitalized.demog.ethnicsmr))
knitr::kable(table.hospitalized.demog.ethnicsmr,
             escape=FALSE, 
             booktabs=TRUE,
             label="ethnicsmr.hosp",
             col.names=c(clean.header(colnames(table.hospitalized.demog.ethnicsmr)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6), 
             caption="Associations of hospitalized or severe disease with ethnicity in those with ethnicity recorded in Scottish Morbidity record") %>%
    add_header_above(c(" "=3, "Univariate"=2, "Multivariate"=2)) %>%
    kable_styling(latex_options="hold_position") %>% 
    column_spec(column=1, width="2.7cm") %>%
    column_spec(column=2, width="2cm")  %>%
    column_spec(column=3, width="1.85cm")  %>%
    column_spec(column=4, width="2.5cm") %>%
    column_spec(column=5, width="1.5cm") %>%
    column_spec(column=6, width="2.5cm") %>%
    column_spec(column=7, width="1.5cm")

```

```{r ethnic5smr, echo=FALSE, warning=FALSE, message=FALSE}

rownames(table.ethnic5smr) <- gsub("_", " ", rownames(table.ethnic5smr))
knitr::kable(table.ethnic5smr,
             escape=FALSE, 
             booktabs=TRUE,
             label="ethnicsmr",
             col.names=c(clean.header(colnames(table.ethnic5smr)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6), 
             caption="Associations of severe disease with ethnicity in those with ethnicity recorded in Scottish Morbidity record") %>%
    add_header_above(c(" "=3, "Univariate"=2, "Multivariate"=2)) %>%
    kable_styling(latex_options="hold_position") %>% 
    column_spec(column=1, width="2.7cm") %>%
    column_spec(column=2, width="1.85cm")  %>%
    column_spec(column=3, width="1.85cm")  %>%
    column_spec(column=4, width="2.5cm") %>%
    column_spec(column=5, width="1.5cm") %>%
    column_spec(column=6, width="2.8cm") %>%
    column_spec(column=7, width="1.5cm")
```

```{r ethnic5smrNRS, echo=FALSE, warning=FALSE, message=FALSE}

rownames(table.ethnic5smrNRS) <- gsub("_", " ", rownames(table.ethnic5smrNRS))
knitr::kable(table.ethnic5smrNRS,
             escape=FALSE, 
             booktabs=TRUE,
             label="ethnicsmrNRS",
             col.names=c(clean.header(colnames(table.ethnic5smrNRS)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6), 
             caption="Associations of severe disease OR a COVID-19 death with ethnicity in those with ethnicity recorded in Scottish Morbidity record") %>%
    add_header_above(c(" "=3, "Univariate"=2, "Multivariate"=2)) %>%
    kable_styling(latex_options="hold_position") %>% 
    column_spec(column=1, width="2.7cm") %>%
    column_spec(column=2, width="1.85cm")  %>%
    column_spec(column=3, width="1.85cm")  %>%
    column_spec(column=4, width="2.5cm") %>%
    column_spec(column=5, width="1.5cm") %>%
    column_spec(column=6, width="2.8cm") %>%
    column_spec(column=7, width="1.5cm")
```

```{r testpositives.ethnic.smr, echo=FALSE, warning=FALSE, message=FALSE}

knitr::kable(testpositives.ethnic.smr, 
           escape=FALSE, 
             booktabs=TRUE,
       caption="\\label{tab:severity.smr}Severity of disease in test-positive cases, by ethnicity assigned in SMR records") %>%
    kable_styling(latex_options="hold_position")

```

\newpage

```{r onomap.testpos, echo=FALSE, warning=FALSE, message=FALSE}

rownames(table.testpositives.demog) <- gsub("_", " ", rownames(table.testpositives.demog))
knitr::kable(table.testpositives.demog,
             escape=FALSE, 
             booktabs=TRUE,
             label="testpos.onomap",
             col.names=c(clean.header(colnames(table.testpositives.demog)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6), 
             caption="Associations of test-positive status with ethnicity based on name classification") %>%
    add_header_above(c(" "=3, "Univariate"=2, "Multivariate"=2)) %>%
    kable_styling(latex_options="hold_position") %>% 
    column_spec(column=1, width="2.7cm") %>%
    column_spec(column=2, width="1.95cm")  %>%
    column_spec(column=3, width="1.85cm")  %>%
    column_spec(column=4, width="2.5cm") %>%
    column_spec(column=5, width="1.5cm") %>%
    column_spec(column=6, width="2.5cm") %>%
    column_spec(column=7, width="1.5cm")

```

```{r onomap.hosp, echo=FALSE, warning=FALSE, message=FALSE}

rownames(table.hospitalized.demog) <- gsub("_", " ", rownames(table.hospitalized.demog))
knitr::kable(table.hospitalized.demog,
             escape=FALSE, 
             booktabs=TRUE,
             label="onomap.hosp",
             col.names=c(clean.header(colnames(table.hospitalized.demog)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6), 
             caption="Associations of hospitalized or severe disease with demographic variables including ethnicity based on name classification") %>%
    add_header_above(c(" "=3, "Univariate"=2, "Multivariate"=2))  %>%
    kable_styling(latex_options="hold_position") %>% 
    column_spec(column=1, width="2.7cm") %>%
    column_spec(column=2, width="2cm")  %>%
    column_spec(column=3, width="1.85cm")  %>%
    column_spec(column=4, width="2.5cm") %>%
    column_spec(column=5, width="1.5cm") %>%
    column_spec(column=6, width="2.5cm") %>%
    column_spec(column=7, width="1.5cm")

```

```{r onomap, echo=FALSE, warning=FALSE, message=FALSE}

rownames(table.severe.demog) <- gsub("_", " ", rownames(table.severe.demog))
knitr::kable(table.severe.demog,
             escape=FALSE, 
             booktabs=TRUE,
             label="onomap",
             col.names=c(clean.header(colnames(table.severe.demog)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6),
             caption="Associations of severe disease with demographic variables including ethnicity based on name classification") %>%
    kable_styling(latex_options="hold_position") %>% 
    column_spec(column=1, width="2cm") %>%
    add_header_above(c(" "=3, "Univariate"=2, "Multivariate"=2)) %>%
    column_spec(column=2, width="1.85cm")  %>%
    column_spec(column=3, width="1.85cm")  %>%
    column_spec(column=4, width="2.5cm") %>%
    column_spec(column=5, width="1.5cm") %>%
    column_spec(column=6, width="2.5cm") %>%
    column_spec(column=7, width="1.5cm")

```

```{r onomapNRS, echo=FALSE, warning=FALSE, message=FALSE}

rownames(table.severe.demogNRS) <- gsub("_", " ", rownames(table.severe.demogNRS))
knitr::kable(table.severe.demogNRS,
             escape=FALSE, 
             booktabs=TRUE,
             label="onomapNRS",
             col.names=c(clean.header(colnames(table.severe.demogNRS)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6),
             caption="Associations of severe disease OR any COVID-19 death with demographic variables including ethnicity based on name classification") %>%
    kable_styling(latex_options="hold_position") %>% 
    column_spec(column=1, width="2cm") %>%
    add_header_above(c(" "=3, "Univariate"=2, "Multivariate"=2)) %>%
    column_spec(column=2, width="1.85cm")  %>%
    column_spec(column=3, width="1.85cm")  %>%
    column_spec(column=4, width="2.5cm") %>%
    column_spec(column=5, width="1.5cm") %>%
    column_spec(column=6, width="2.5cm") %>%
    column_spec(column=7, width="1.5cm")

```

```{r testpositives.ethnic, echo=FALSE, warning=FALSE, message=FALSE}

knitr::kable(testpositives.ethnic, 
           escape=FALSE, 
             booktabs=TRUE,
       caption="\\label{tab:severity}Severity of disease in test-positive cases, by ethnicity assigned from ONOMAP name classification") %>%
    kable_styling(latex_options="hold_position")

```

```{r testpositives.carehome, echo=FALSE, warning=FALSE, message=FALSE}

knitr::kable(testpositives.carehome, 
           escape=FALSE, 
             booktabs=TRUE,
       caption="\\label{tab:carehome}Care home status in test-positive cases, by ethnicity assigned from ONOMAP name classification") %>%
    kable_styling(latex_options="hold_position")

```

\newpage

```{r testpositives.healthboard, echo=FALSE, warning=FALSE, message=FALSE}

knitr::kable(testpositives.healthboard, 
             escape=FALSE, 
             booktabs=TRUE,
             caption="\\label{tab:board}Health board of residence in test-positive cases, by ethnicity assigned from ONOMAP name classification") %>%
    kable_styling(latex_options="hold_position")

```