### 2020-06-19 casecontrol_script.Rmd
fixed miscoding of broad definition of fatal cases

---
title: 'REACT-SCOT analysis script report'
author: Paul McKeigue

header-includes: |
  \let\origquote\quote
  \def\quote{\origquote\itshape}
  \usepackage{graphicx}
  \usepackage{longtable}
  \usepackage{booktabs}
  \usepackage{float}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{longtable}
  \usepackage{lscape}
output: 
  pdf_document

always_allow_html: true
urlcolor: blue
---

```{r functions, echo=FALSE}
library(knitr)
library(kableExtra)
library(bookdown)

# rmarkdown::render("casecontrol_script.Rmd", output_file="casecontrol_script.pdf")

options(knitr.table.format = "latex") # adds tab: prefix to labels
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)

source("helperfunctions.R")

```

\newpage

\newpage

# Tables

```{r death28, echo=FALSE, warning=FALSE, message=FALSE}
options(knitr.kable.NA = '')

knitr::kable(with(cc.all[cc.all$CASE==1, ],
                  paste.colpercent(table(deathwithin28, unitcategory))), 
             escape=FALSE, 
             booktabs=TRUE,
             row.names=FALSE,
             caption="Death within 28 days by hospitalization group") %>%
    kable_styling(latex_options="hold_position")

```

```{r casegroup, echo=FALSE, warning=FALSE, message=FALSE}

knitr::kable(table(cc.all$CASE, cc.all$group, exclude=NULL),
             escape=FALSE, 
             booktabs=TRUE,
             row.names=FALSE,
             caption="Case groups") %>%
    kable_styling(latex_options="hold_position")

```

```{r NRS.cases, echo=FALSE, warning=FALSE, message=FALSE}

knitr::kable(with(subset(cc.all, CASE==1 & nrs_covid_case==0), table(criticalcare, deathwithin28, exclude=NULL)),
             caption="CASE x NRS COVID case")

knitr::kable(with(subset(cc.all, CASE==1), table(criticalcare, nrs_covid_case, exclude=NULL)),
             caption="Critical care x NRS COVID case")

knitr::kable(with(subset(cc.all, CASE==1), table(nrs_covid_case, deathwithin28, exclude=NULL)),
             caption="NRS covid case x death within 28 days")

```