### 2020-07-01 Covid_ethnicity_Scotland_reportPartB.Rmd
Updated to remove disaggregation of Black ethnicity in onomap tables.

---
title: "Briefing paper on Ethnicity and COVID-19 status in Scotland"

author: Modelling & Research Cell, Public Health Scotland

date: "`r gsub('^0', '', format(Sys.time(), '%d %B %Y'))`"

header-includes: |
  \let\origquote\quote
  \def\quote{\origquote\itshape}
  \usepackage{graphicx}
  \usepackage{longtable}
  \usepackage{booktabs}
  \usepackage{float}
  \usepackage{array}
  \usepackage{longtable}
  \usepackage{lscape}

output:
  pdf_document:
  keep_tex: true
  latex_engine: lualatex

# rmarkdown::render("Covid_ethnicity_Scotland.Rmd")

---

```{r settings, echo = FALSE, results='asis', warning=FALSE, error=FALSE}

# GLobal R options for this document
options(knitr.table.format="latex")
options(knitr.kable.NA='')

```

# Part B Tables
In this paper we attempt carry out a more detailed disaggregation of ethnicity presented in the "Briefing paper on Ethnicity and COVID-19 status in Scotland".

The following tables further disaggregate the ethnicity distributions - the coding to these disaggregated groups is as follows;

SMR recodes:-
1L - White Polish.
3B/3C/3F/3H - Pakistani/Bangladeshi.
3A/3G - Indian.
4A/4E/5C - Caribbean.
4B/4D/4Y - African.

ONOMAP "OnolyticsType" recodes:-
White - White and White Polish.
Black - Other
Bangladeshi - Muslim South Asian.
Muslim Indian - Muslim South Asian.
Pakistani - Muslim South Asian.
Pakistani Kashmir - Muslim South Asian.
Muslim - Muslim South Asian.

```{r ethnicB, echo=FALSE, warning=FALSE, message=FALSE}

knitr::kable(table.ethnicB, 
             escape=FALSE, 
             booktabs=TRUE,
             caption="\\label{tab:crosstab}Comparison of ethnicity assigned by ONOMAP name classification (rows) with ethnicity assigned in SMR  records (columns)") %>%
  add_header_above(c(" "=1, "SMR records"= 10)) %>%
  pack_rows("ONOMAP", 1, 5) %>%
  landscape() %>%
  kable_styling(latex_options="hold_position", font_size=8)

```

```{r ethnicsmr.testposB, echo=FALSE, warning=FALSE, message=FALSE}

rownames(table.testpositives.demog.ethnicsmrB) <-
  gsub("_", " ", rownames(table.testpositives.demog.ethnicsmrB))
knitr::kable(table.testpositives.demog.ethnicsmrB,
             escape=FALSE, 
             booktabs=TRUE,
             label="testpos.ethnicsmr",
             col.names=c(clean.header(colnames(table.testpositives.demog.ethnicsmrB)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6), 
             caption="Associations of test-positive status with ethnicity in those with ethnicity recorded in Scottish Morbidity record") %>%
  kable_styling(latex_options="hold_position") %>% 
  column_spec(column=1, width="2.7cm")  %>%
  column_spec(column=2, width="2cm")  %>%
  column_spec(column=3, width="1.85cm")  %>%
  column_spec(column=4, width="2.5cm") %>%
  column_spec(column=5, width="1.5cm") %>%
  column_spec(column=6, width="2.5cm") %>%
  column_spec(column=7, width="1.5cm")
```

```{r ethnicsmr.hospB, echo=FALSE, warning=FALSE, message=FALSE}

rownames(table.hospitalized.demog.ethnicsmrB) <-
  gsub("_", " ", rownames(table.hospitalized.demog.ethnicsmrB))
knitr::kable(table.hospitalized.demog.ethnicsmrB,
             escape=FALSE, 
             booktabs=TRUE,
             label="ethnicsmr.hosp",
             col.names=c(clean.header(colnames(table.hospitalized.demog.ethnicsmrB)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6), 
             caption="Associations of hospitalized or severe disease among test positives with ethnicity in those with ethnicity recorded in Scottish Morbidity record") %>%
  kable_styling(latex_options="hold_position") %>% 
  column_spec(column=1, width="2.7cm") %>%
  column_spec(column=2, width="2cm")  %>%
  column_spec(column=3, width="1.85cm")  %>%
  column_spec(column=4, width="2.5cm") %>%
  column_spec(column=5, width="1.5cm") %>%
  column_spec(column=6, width="2.5cm") %>%
  column_spec(column=7, width="1.5cm")

```

```{r ethnic5smrB, echo=FALSE, warning=FALSE, message=FALSE}

rownames(table.ethnic9smrB) <- gsub("_", " ", rownames(table.ethnic9smrB))
knitr::kable(table.ethnic9smrB,
             escape=FALSE, 
             booktabs=TRUE,
             label="ethnicsmr",
             col.names=c(clean.header(colnames(table.ethnic9smrB)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6), 
             caption="Associations of severe disease among test positives with ethnicity in those with ethnicity recorded in Scottish Morbidity record") %>%
  kable_styling(latex_options="hold_position") %>% 
  column_spec(column=1, width="2.7cm") %>%
  column_spec(column=2, width="1.85cm")  %>%
  column_spec(column=3, width="1.85cm")  %>%
  column_spec(column=4, width="2.5cm") %>%
  column_spec(column=5, width="1.5cm") %>%
  column_spec(column=6, width="2.8cm") %>%
  column_spec(column=7, width="1.5cm")
```

```{r ethnic9smrNRSB, echo=FALSE, warning=FALSE, message=FALSE}

rownames(table.ethnic9smrNRSB) <- gsub("_", " ", rownames(table.ethnic9smrNRSB))
knitr::kable(table.ethnic9smrNRSB,
             escape=FALSE, 
             booktabs=TRUE,
             label="ethnicsmrNRSB",
             col.names=c(clean.header(colnames(table.ethnic9smrNRSB)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6), 
             caption="Associations of severe disease OR a COVID-19 death with ethnicity in those with ethnicity recorded in Scottish Morbidity record") %>%
    add_header_above(c(" "=3, "Univariate"=2, "Multivariate"=2)) %>%
    kable_styling(latex_options="hold_position") %>% 
    column_spec(column=1, width="2.7cm") %>%
    column_spec(column=2, width="1.85cm")  %>%
    column_spec(column=3, width="1.85cm")  %>%
    column_spec(column=4, width="2.5cm") %>%
    column_spec(column=5, width="1.5cm") %>%
    column_spec(column=6, width="2.8cm") %>%
    column_spec(column=7, width="1.5cm")
```

```{r testpositives.ethnic.smrB, echo=FALSE, warning=FALSE, message=FALSE}

knitr::kable(testpositives.ethnic.smrB, 
             escape=FALSE, 
             booktabs=TRUE,
             caption="\\label{tab:severity.smr}Severity of disease in test-positive cases, by ethnicity assigned in SMR records") %>%
  kable_styling(latex_options="hold_position")

```

\newpage

```{r onomap.testposB, echo=FALSE, warning=FALSE, message=FALSE}

rownames(table.testpositives.demogB) <- gsub("_", " ", rownames(table.testpositives.demogB))
knitr::kable(table.testpositives.demogB,
             escape=FALSE, 
             booktabs=TRUE,
             label="testpos.onomap",
             col.names=c(clean.header(colnames(table.testpositives.demogB)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6), 
             caption="Associations of test-positive status with ethnicity based on name classification") %>%
  add_header_above(c(" "=3, "Univariate"=2, "Multivariate"=2)) %>%
  kable_styling(latex_options="hold_position") %>% 
  column_spec(column=1, width="2.7cm") %>%
  column_spec(column=2, width="1.95cm")  %>%
  column_spec(column=3, width="1.85cm")  %>%
  column_spec(column=4, width="2.5cm") %>%
  column_spec(column=5, width="1.5cm") %>%
  column_spec(column=6, width="2.5cm") %>%
  column_spec(column=7, width="1.5cm")

```

```{r onomap.hospB, echo=FALSE, warning=FALSE, message=FALSE}

rownames(table.hospitalized.demogB6) <- gsub("_", " ", rownames(table.hospitalized.demogB6))
knitr::kable(table.hospitalized.demogB6,
             escape=FALSE, 
             booktabs=TRUE,
             label="onomap.hosp",
             col.names=c(clean.header(colnames(table.hospitalized.demogB6)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6), 
             caption="Associations of hospitalized or severe disease among test positives with demographic variables including ethnicity based on name classification") %>%
  add_header_above(c(" "=3, "Univariate"=2, "Multivariate"=2))  %>%
  kable_styling(latex_options="hold_position") %>% 
  column_spec(column=1, width="2.7cm") %>%
  column_spec(column=2, width="2cm")  %>%
  column_spec(column=3, width="1.85cm")  %>%
  column_spec(column=4, width="2.5cm") %>%
  column_spec(column=5, width="1.5cm") %>%
  column_spec(column=6, width="2.5cm") %>%
  column_spec(column=7, width="1.5cm")

```

```{r onomapB, echo=FALSE, warning=FALSE, message=FALSE}

#rownames(table.severe.demogB) <- gsub("_", " ", rownames(table.severe.demogB))
#knitr::kable(table.severe.demogB,
#             escape=FALSE, 
#             booktabs=TRUE,
#             label="onomapB",
#             col.names=c(clean.header(colnames(table.severe.demogB)[1:2]),
#                         "Rate ratio (95\\% CI)", "p-value",
#                         "Rate ratio (95\\% CI)", "p-value"),
#             align=rep("c", 6),
#             caption="Associations of severe disease with demographic variables including ethnicity based on name classification") %>%
#  kable_styling(latex_options="hold_position") %>% 
#  column_spec(column=1, width="2cm") %>%
#  add_header_above(c(" "=3, "Univariate"=2, "Multivariate"=2)) %>%
#  column_spec(column=2, width="1.85cm")  %>%
#  column_spec(column=3, width="1.85cm")  %>%
#  column_spec(column=4, width="2.5cm") %>%
#  column_spec(column=5, width="1.5cm") %>%
#  column_spec(column=6, width="2.5cm") %>%
#  column_spec(column=7, width="1.5cm")

```

```{r onomapB6, echo=FALSE, warning=FALSE, message=FALSE}

rownames(table.severe.demogB6) <- gsub("_", " ", rownames(table.severe.demogB6))
knitr::kable(table.severe.demogB6,
             escape=FALSE, 
             booktabs=TRUE,
             label="onomapB6",
             col.names=c(clean.header(colnames(table.severe.demogB6)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6),
             caption="Associations of severe disease among test positives with demographic variables including ethnicity based on name classification") %>%
  kable_styling(latex_options="hold_position") %>% 
  column_spec(column=1, width="2cm") %>%
  add_header_above(c(" "=3, "Univariate"=2, "Multivariate"=2)) %>%
  column_spec(column=2, width="1.85cm")  %>%
  column_spec(column=3, width="1.85cm")  %>%
  column_spec(column=4, width="2.5cm") %>%
  column_spec(column=5, width="1.5cm") %>%
  column_spec(column=6, width="2.5cm") %>%
  column_spec(column=7, width="1.5cm")

```

```{r onomapNRSB, echo=FALSE, warning=FALSE, message=FALSE}

#rownames(table.severe.demogNRSB) <- gsub("_", " ", rownames(table.severe.demogNRSB))
#knitr::kable(table.severe.demogNRSB,
#             escape=FALSE, 
#             booktabs=TRUE,
#             label="onomapNRSB",
#             col.names=c(clean.header(colnames(table.severe.demogNRSB)[1:2]),
#                         "Rate ratio (95\\% CI)", "p-value",
#                         "Rate ratio (95\\% CI)", "p-value"),
#             align=rep("c", 6),
#             caption="Associations of severe disease OR any COVID-19 death with demographic variables including ethnicity based on name #classification") %>%
#    kable_styling(latex_options="hold_position") %>% 
#    column_spec(column=1, width="2cm") %>%
#    add_header_above(c(" "=3, "Univariate"=2, "Multivariate"=2)) %>%
#    column_spec(column=2, width="1.85cm")  %>%
#    column_spec(column=3, width="1.85cm")  %>%
#    column_spec(column=4, width="2.5cm") %>%
#    column_spec(column=5, width="1.5cm") %>%
#    column_spec(column=6, width="2.5cm") %>%
#    column_spec(column=7, width="1.5cm")

```

```{r onomapNRSB6, echo=FALSE, warning=FALSE, message=FALSE}

rownames(table.severe.demogNRSB6) <- gsub("_", " ", rownames(table.severe.demogNRSB6))
knitr::kable(table.severe.demogNRSB6,
             escape=FALSE, 
             booktabs=TRUE,
             label="onomapNRSB6",
             col.names=c(clean.header(colnames(table.severe.demogNRSB6)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6),
             caption="Associations of severe disease OR any COVID-19 death with demographic variables including ethnicity based on name classification") %>%
    kable_styling(latex_options="hold_position") %>% 
    column_spec(column=1, width="2cm") %>%
    add_header_above(c(" "=3, "Univariate"=2, "Multivariate"=2)) %>%
    column_spec(column=2, width="1.85cm")  %>%
    column_spec(column=3, width="1.85cm")  %>%
    column_spec(column=4, width="2.5cm") %>%
    column_spec(column=5, width="1.5cm") %>%
    column_spec(column=6, width="2.5cm") %>%
    column_spec(column=7, width="1.5cm")

```

```{r testpositives.ethnicB, echo=FALSE, warning=FALSE, message=FALSE}

knitr::kable(testpositives.ethnicB, 
             escape=FALSE, 
             booktabs=TRUE,
             caption="\\label{tab:severity}Severity of disease in test-positive cases, by ethnicity assigned from ONOMAP name classification") %>%
  kable_styling(latex_options="hold_position")

```

```{r testpositives.carehomeB, echo=FALSE, warning=FALSE, message=FALSE}

knitr::kable(testpositives.carehomeB, 
             escape=FALSE, 
             booktabs=TRUE,
             caption="\\label{tab:carehome}Care home status in test-positive cases, by ethnicity assigned from ONOMAP name classification") %>%
  kable_styling(latex_options="hold_position")

```

\newpage

```{r testpositives.healthboardB, echo=FALSE, warning=FALSE, message=FALSE}

knitr::kable(testpositives.healthboardB, 
             escape=FALSE, 
             booktabs=TRUE,
             caption="\\label{tab:board}Health board of residence in test-positive cases, by ethnicity assigned from ONOMAP name classification") %>%
  landscape() %>%
  kable_styling(latex_options="hold_position")

```