### 2021-01-09 beamerkable.Rmd

---
title: "beamer kable" 
author: PM 
output: 
  beamer_presentation:
    slide_level: 1
    includes:
      in_header: /mnt/phs/beamer_header.tex
    keep_tex: yes
---

# Title slide

# Table with kable

```{r kable, echo=FALSE, warning=FALSE, message=FALSE}

library(knitr)
library(kableExtra)

options(knitr.table.format = "latex") # adds tab: prefix to labels

knitr::kable(mtcars[1:4, 1:4], 
             escape=FALSE,
             booktabs=TRUE,
	label="regressionseverecasesnocare",
	caption="Rate ratios for severe COVID-19 in those not resident in care homes" 
	) %>% 
    add_header_above(c(" "=2, "Univariate"=2))

```
# more stuff

           format="latex", 
             escape=FALSE,
             booktabs=TRUE,
             label="regressionseverecasesnocare",
             caption="Rate ratios for severe COVID-19 in those not resident in care homes. ", 
             col.names=c(clean.header(colnames(table.shielded.severecases.nocare)[1:2]),
 #                        "Rate ratio (95\\% CI)", "\\ensuremath{p}-value",
                         "Rate ratio (95\\% CI)", "\\ensuremath{p}-value")) %>%
#    add_header_above(c(" "=3, "Univariate"=2, "Adjusted"=2)) %>%
    column_spec(column=1, width="3cm") %>%
    column_spec(column=2, width="2cm") %>%
    column_spec(column=3, width="2cm") %>%
#    column_spec(column=4, width="2.5cm") %>%
#    column_spec(column=5, width="2cm") %>%
#    column_spec(column=6, width="2.5cm") %>%
#    column_spec(column=7, width="2cm") %>%
    kableExtra::group_rows("Shielding eligibility category", 3, 8) %>%
#    kableExtra::group_rows("School-age children in household", 10, 12) %>%
#    kableExtra::group_rows("Adults in household", 13, 16) %>%
#    kableExtra::group_rows("SIMD quintile", 18, 22) %>%
#    kableExtra::group_rows("Occupation", 23, 26) %>%
    kable_styling(latex_options=c("scale_down", "HOLD_position")) %>%
    add_footnote(label=c("Percentages are column percentages for each variable",
                         "PF, patient-facing"), notation="none")