### 2021-02-10 beamerkable.Rmd
updated shielding markdown script before circulation

---
title: "Morbidity and transmission factors associated with variant B1.1.7" 
author: "Paul McKeigue and Helen Colhoun" 
output: 
  beamer_presentation:
    slide_level: 1
    includes:
      in_header: ../beamer_header.tex
    keep_tex: yes
---

# 28-day fatality by S gene dropout status
\scriptsize
\setbeamertemplate{caption}{\raggedright\insertcaption\par}

```{r kable, echo=FALSE, warning=FALSE, message=FALSE}

library(knitr)
library(kableExtra)

options(knitr.table.format = "latex") # adds tab: prefix to labels

knitr::kable(table.death.coxph[3:5, c(1:2, 5:6)],
             escape=FALSE, 
             booktabs=TRUE,
			 label="coxfatal",
             row.names=TRUE,
             align=c("r", "r", "r", "l"), 
             col.names=c(clean.header(colnames(table.death.coxph)[1:2]),
                        "Rate ratio (95\\% CI)", "\\ensuremath{p}-value"), 
             caption="Death within 28 days of testing positive: Cox regression on S gene dropout and covariates") %>%
    column_spec(column=1, width="2cm") %>%
    column_spec(column=2, width="2cm") %>%
    column_spec(column=3, width="1.8cm") %>%
    column_spec(column=4, width="2.2cm") %>%
    column_spec(column=5, width="1.5cm") %>%
    add_footnote(label=c("Adjusted for age and sex",
                         "Stratified by days since testing positive and calendar week"),
                 notation="none")

```
# Test-positive cases: severity and cycle threshold

```{r plotvariantbydate, echo=FALSE, warning=FALSE, message=FALSE, fig.pos = "H", fig.asp=0.8}

theme_set(theme_gray(base_size = 18))

gA <- ggplotGrob(p.variantbydate)
gB <- ggplotGrob(p.mediansbydate)
grid::grid.newpage()
grid::grid.draw(rbind(gA, gB))

```

# Severe cases by source

```{r plotvariantbysource, echo=FALSE, warning=FALSE, message=FALSE, fig.pos = "H", fig.asp=0.75}

theme_set(theme_gray(base_size = 18))
p.bysource

```