### 2021-02-10 shielding.Rmd

---
title: "Relation of severe COVID-19 in Scotland to transmission-related factors in individuals with and without risk conditions eligible for shielding support: REACT-SCOT case-control study"
author:
  - name: Paul M McKeigue\textsuperscript{\getAff{Usher}}
    affiliation: HPS
  - name: David McAllister\textsuperscript{\getAff{Glasgow}}  
    affiliation: HPS 
  - name: David Caldwell 
    affiliation: HPS
  - name: Ciara Gribben
    affiliation: HPS
  - name: Jen Bishop
    affiliation: HPS
  - name: Matthew Armstrong
    affiliation: HPS
  - name: Joke Delvaux
    affiliation: HPS
  - name: Sam Colville
    affiliation: HPS
  - name: Sharon Hutchinson\textsuperscript{\getAff{GCU}} 
    affiliation: HPS
  - name: Chris Robertson\textsuperscript{\getAff{Strathclyde}}   
    affiliation: HPS
  - name: Nazir Lone
    affiliation: Usher
  - name: Jim McMenamin 
    affiliation: HPS
  - name: David Goldberg
    affiliation: HPS
  - name: Helen M Colhoun\textsuperscript{\getAff{IGMM}}
    email: helen.colhoun@igmm.ed.ac.uk (+44 7791 652665)
    affiliation: HPS
address: 
- code: Usher
  address: Usher Institute, College of Medicine and Veterinary Medicine, University of Edinburgh, Teviot Place, Edinburgh EH8 9AG, Scotland. PM - Professor of Genetic Epidemiology and Statistical Genetics. NL - Clinical Senior Lecturer in Critical Care  
- code: HPS
  address: Public Health Scotland, Meridian Court, 5 Cadogan Street, Glasgow G2 6QE
- code: Glasgow
  address: Institute of Health and Wellbeing, University of Glasgow, 1 Lilybank Gardens, Glasgow G12 8RZ. DM - Senior Clinical Lecturer in Public Health 
- code: GCU
  address: School of Health and Life Sciences, Glasgow Caledonian University. SH - Professor of Epidemiology and Population Health 
- code: Strathclyde
  address: Department of Mathematics and Statistics, University of Strathclyde, 16 Richmond Street, Glasgow G1 1XQ. CR - Professor of Public Health Epidemiology
- code: IGMM
  address: Institute of Genetics and Molecular Medicine, College of Medicine and Veterinary Medicine, University of Edinburgh, Western General Hospital Campus, Crewe Road, Edinburgh EH4 2XUC, Scotland. HC - AXA Chair in Medical Informatics and Epidemiology

header-includes:
  \usepackage{newunicodechar}
  \let\origquote\quote
  \def\quote{\origquote\itshape}
  \usepackage{graphicx}
  \usepackage{geometry}
  \usepackage{longtable}
  \usepackage{booktabs}
  \usepackage{float}
  \floatplacement{figure}{H}
  \usepackage{array}
  \usepackage{threeparttable}
  \usepackage{longtable}
  \usepackage{lscape}
  \usepackage{pdfpages}
output:
  rticles::nejm_article:
    includes:
      in_header: ./preamble.tex
    latex_engine: lualatex
    keep_tex: true
    fig_caption: yes
#    number_sections: false
    md_extensions: -autolink_bare_uris
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
bibliography: ./covid.bib
csl: ./biomed-central.csl
always_allow_html: true
urlcolor: blue
linenumbers: false
linkcolor: cyan

---

```{r functions, echo=FALSE}
library(knitr)
library(kableExtra)

options(knitr.kable.NA = '.')
options(knitr.table.format = "latex") # adds tab: prefix to labels
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
knitr::opts_chunk$set(fig.width=6, fig.pos = "H", out.extra = "")
source("helperfunctions.R")

# rmarkdown::render("shielding.Rmd", output_file="shielding.pdf")

dateformat <- function(x) { # format date and strip leading zero
     gsub("^0", "", format(x, "%d %B %Y"))
}

format.estcipv <- function(x.ci, x.pval) {
    x <- gsub(" \\(", " \\(95\\% CI ", as.character(x.ci))
    x <- gsub(", ", " to ", x)
    x <- gsub("\\)", paste0(", _p_=\\", x.pval, "\\)"), x)
    x <- gsub("times", "\\\\times", x)
    return(x)
}

peakdate <- casedates.allgr[which.max(avdaily), date]
maxdeaths <- casedates.allgr[which.max(avdaily), avdaily]
deaths.quartermax.date <- casedates.allgr[date > peakdate & avdaily < 0.25 * maxdeaths,
                                 date][1]

peakdate.elig <- casedates.byelig[listedgr3=="Eligible for shielding"][which.max(avdaily), date]
maxdeaths.elig <- casedates.byelig[listedgr3=="Eligible for shielding"][which.max(avdaily), avdaily]
deaths.elig.quartermax.date <- casedates.byelig[listedgr3=="Eligible for shielding" &
                                         date > peakdate.elig &
                                         avdaily < 0.25 * maxdeaths.elig,
                                         date][1]

r <- as.numeric(gsub(" .*", "", table.shieldedonly.severecases.nocare["Recent hospital visit/stay", 5]));

pc <- gsub(".*\\(", "", table.shieldedonly.severecases.nocare["Recent hospital visit/stay", 2])
pc <- 0.01 * as.numeric(gsub("\\\\.*", "", pc))

```

 \newpage

# Abstract

* _Objectives_ - To investigate the relation of transmission-related factors-- occupation, household composition and recent hospital exposure -- to severe COVID-19 in  clinically vulnerable individuals eligible for shielding and in the general population.

* _Design_ - Matched case-control study (REACT-SCOT).

* _Setting_ - population of Scotland from 1 March 2020 to `r dateformat(lastdate.specimen)`.

* _Participants_ -  All `r nrow(cc.all[CASE==1])` diagnosed cases of COVID-19 and `r nrow(cc.all[CASE==0])` controls matched for age, sex and primary care practice, linked with all `r nrow(shielded.full)` individuals identified as eligible for shielding by Public Health Scotland.

* _Main outcome measure_ - severe COVID-19, defined as cases that entered critical care or were  fatal.

* _Results_ - With those without risk conditions as reference category, the univariate rate ratio for severe COVID-19 was
`r format.estcipv(table.listedgr3.severecases.nocare["Moderate risk condition", 3], table.listedgr3.severecases.nocare["Moderate risk condition", 4])` 
in those with moderate risk conditions and
`r format.estcipv(table.listedgr3.severecases.nocare["Eligible for shielding", 3], table.listedgr3.severecases.nocare["Eligible for shielding", 4])` 
in those eligible for shielding.  The highest risk for severe disease was in solid organ transplant recipients: rate ratio `r format.estcipv(table.shielded.severecases.nocare["Solid organ transplant", 3], table.shielded.severecases.nocare["Solid organ transplant", 4])` .   In all risk groups severe COVID-19 was strongly associated with recent hospital exposure: rate ratio `r format.estcipv(table.shielded.severecases.nocare["Recent hospital visit/stay", 3], table.shielded.severecases.nocare["Recent hospital visit/stay", 4])`  overall. The fraction of severe cases attributable to recent hospital exposure peaked at  `r round(100 * coeffs.hosp.timewindow[date==as.Date("2020-05-01"), popattr.frac])`% at the beginning of May 2020, and again at `r round(100 * coeffs.hosp.timewindow[date==as.Date("2020-12-01"), popattr.frac])`%  at the beginning of December 2020. Risk of severe COVID-19 increased with the number of adults in the household but decreased with the number of school-age children in the household.  The rate ratio associated with two or more adults (with single-adult households as reference category), was   `r format.estcipv(table.listedgr3.severecases.nocare["adultsgt1", 5], table.listedgr3.severecases.nocare["adultsgt1", 6])` and the rate ratio associated with one or more school-age children was `r format.estcipv(table.listedgr3.severecases.nocare.anyschoolage["hh.schoolage.any", 5], table.listedgr3.severecases.nocare.anyschoolage["hh.schoolage.any", 6])`.

* _Conclusions_ -  Mitigating the impact of the epidemic requires control of transmission in hospital settings.  Shielding of vulnerable individuals until they have been vaccinated requires protection from transmission from other adults in the same household.  For solid organ transplant recipients, who are at very high risk and in whom the efficacy of vaccines is uncertain, vaccination should be offered to household contacts.

# Background  
The SARS-CoV-2 pandemic reached Scotland in early 2020 with the first positive laboratory test recorded on 1 March 2020. Population-wide interventions included advice that symptomatic individuals should self-isolate, banning of mass gatherings, closure of schools and finally a population-wide lockdown on 23 March 2020.  Although no systematic studies of risk factors were available at that time, public health agencies issued lists of "moderate risk" conditions [@nhs_whos_2020] and "diseases and conditions considered to be very high risk" [@kanani_caring_2020].  Those with "very high risk" conditions were designated as clinically extremely vulnerable and eligible for shielding.

The shielding programme issued individual letters to eligible persons advising on the need for social distancing, recommending strict isolation and giving information on the provided support which included a national opt-in scheme of free food delivery, home delivery of medication and priority access to supermarket delivery slots. Whether the programme had an impact on the infection rates in those shielded is unclear.

The initial objective of this study was to investigate  whether the shielding letter and programme had any demonstrable impact on the risk of COVID-19 in those eligible for shielding  versus those with moderate risk or the rest of the population.  For this we hypothesized that  a fall in  the risk ratio  for those eligible for  shielding  after the receipt of shielding letters  would be evidence of impact. Our second objective was to understand the role of  transmission-related factors: household composition, recent hospital exposure including out-patient consultations, and exposure related to occupations including health care [@shah_risk_2020] and teaching [@mcallister_report_2020]  on COVID-19 among those eligible for shielding.    As the relevance of these transmission-related factors among the shielded  became clear, the objectives were broadened to investigation of the relation of severe COVID-19 to these transmission-related factors in  the remainder of the population also.

# Methods
In brief; we linked a dataset of all those eligible for shielding  nationally to an existing case-control dataset (REACT-SCOT) conducted by PHS that includes all COVID-19 cases in Scotland and matched population controls.[@mckeigue_rapid_2020-1]   This case control dataset is refreshed every few weeks and  through  linkage to health and other records already contained extensive covariate data on a wide range of risk factors and settings for COVID-19 and derived variables for   "moderate risk conditions". Thus all cases and controls were  classifiable into three categories: no risk condition, moderate risk condition only, eligible for shielding.  The rate ratio of COVID-19 associated with each risk category was estimated  as the conditional odds ratio  from the case-control study as described below.

## Ascertainment of eligibility for shielding 
We used the dataset of those eligible for shielding that was created by PHS for the purposes of issuing letters.   This dataset comprised a list of  patient identifiers and the shielding condition and the date the shielding letter was issued. The list had been generated  by querying health-care information systems including hospital discharge records, prescription encashments, regional cancer chemotherapy databases, blood and transplant registries,  for evidence of a list of agreed diseases and conditions  and supplemented by data submitted  to PHS following direct requests to clinicians in relevant specialties [@health_protection_scotland_search_2020].

The categories designated as eligible for shielding were as listed below [@health_protection_scotland_search_2020; @kanani_caring_2020]:

1. Solid organ transplant recipients
2. People with cancer of the blood or bone marrow at any stage of treatment, or people with cancer receiving treatments that affect the immune system 
3. People with severe respiratory conditions including cystic fibrosis, severe asthma and severe chronic obstructive airway disease, on home oxygen, severe bronchiectasis, pulmonary hypertension)
4. People with rare diseases that increase the risk of infections such as severe combined immunodeficiency, homozygous sickle cell disease. 
5. People on immunosuppression therapies sufficient to increase risk of infection
6. Women who were pregnant with heart disease
7. Other patients identified by clinicians as requiring shielding advice. 
8. Those on renal dialysis and those who had had a splenectomy (included from 11 May 2020)

The dataset of  those eligible for shielding process  has been regularly refreshed.  Here we used the dataset comprising  `r nrow(shielded.full)` individuals identified up to `r dateformat(lastdate.specimen)`. The first batch of  shielding letters was sent on `r dateformat(batches$Date.Sent)[1]`.   Further batches were issued on a weekly basis and the programme was paused on  1 August 2020. On 25 November 2020 a further letter was issued with "extra protection level advice for people at highest risk" based on the current protection level for the population level in that area.

## Ascertainment of cases and sampling of controls
Case ascertainment for the REACT-SCOT study has been described in detail elsewhere [@mckeigue_rapid_2020-1].  Case ascertainment was based on querying the following national-level databases: Electronic Communication of Surveillance in Scotland (ECOSS) that captures virology testing in all NHS laboratories, National Records of Scotland (NRS) death registrations,  RAPID which is a daily update of hospitalisations, and Scottish Morbidity Record 01 (SMR01) which records general hospital discharges including day cases and is ICD-10 coded. All these databases use the Community Health Index (CHI) number as identifier.  The CHI database includes age, sex, postcode and care home status, and can be queried to extract numbers of adults and children in the household.

Cases of COVID-19 were defined as those with a positive  nucleic acid test for SARS-CoV-2 in ECOSS, a hospital discharge diagnosis of COVID-19 in SMR01, or a death registration with mention of COVID-19 anywhere on the death certificate. The presentation date was assigned as the date of the first positive test for those ascertained through testing, as seven days before the admission date for those without a positive test result ascertained through hospital discharge records, and as fourteen days before the date of death for those without a positive test result ascertained through death certificates. Databases were queried from 1 March 2020  (date of the first diagnosed case of COVID-19 in Scotland) up to `r dateformat(maxdate.specimen)` for test results and `r dateformat(maxdate.death)` for deaths.

Severe COVID-19 was defined by entry to critical care within 21 days of presentation date, or fatal outcome. Entry to critical care units -- intensive care, high dependency or combined units -- was obtained by linkage to the Scottish Intensive Care Society and Audit Group (SICSAG) database. Fatal outcome was defined as death at any time with COVID-19 coded as underlying cause, death from any cause within 28 days of testing positive or death within 28 days of presentation date for cases ascertained through discharge records.  This definition of severe cases was chosen to ensure that case ascertainment would not be affected by triage of those assessed as unlikely to benefit from critical care, or by the scale-up of SARS-CoV-2 testing.

For each case of COVID-19, up to 10  community controls matched for sex, one-year age band and primary care practice were selected from the CHI database, and assigned the same presentation date as the case. With this incidence density sampling design, it is possible and correct for an individual to appear more than once as a control and subsequently as a case.

## Linkage of cases and controls to demographic and morbidity data
Linkage of cases and controls to demographic and morbidity data and the associations of these factors with severe COVID-19 have been described in detail elsewhere [@mckeigue_rapid_2020-1].  Cases and controls were linked to hospital discharge ICD-10 codes over the last five years in SMR01, to British National Formulary codes of dispensed prescriptions in the 240 days before presentation date in the Prescribing Information System (PIS) and to the national register of diabetes.  We used these linked datasets to assign a list of "moderate risk conditions" for COVID-19 designated by public health agencies [@nhs_whos_2020]: diabetes, heart disease, asthma or chronic airway disease, chronic kidney disease, disabling neurological conditions, and immune deficiency or suppression.  The codes used are as described previously  [@mckeigue_rapid_2020-1].  Three risk groups were defined: no risk condition, moderate risk condition but ineligible for shielding, and eligible for shielding.  Socioeconomic status was assigned as the Scottish Index of Multiple Deprivation (SIMD) score which is based on linkage of postcodes to Census data [@fraser_scottish_2020].  Quintile 5 is the least deprived.

# Transmission-related risk factors 
Through linkage of the CHI database to the Unique Property Reference Number it was possible to calculate the numbers of adults and children in each household and to refine the care home residence  assignation contained in the CHI database.  Care home residence was assigned using the field in the CHI database, augmented by coding as care home residents the `r care.home.reclassified[1]` individuals aged over 70 in households with 10 or more adults to give a total of `r cc.all[care.home=="Care/nursing home", .N]` care home residents out of the `r cc.all[, .N]` cases and controls.  Linkage to occupational status for health-care workers and teachers was undertaken as described elsewhere [@shah_risk_2020; @mcallister_report_2020].

We used the Scottish Morbidity Records SMR01 and SMR00 (outpatient attendance) to examine possible associations with exposure to health-care facilities. We defined the variable "recent hospital exposure" as any hospital in-patient stay or face-to-face out-patient consultation from 5 to 14 days before presentation date. Restriction of hospital exposure to this time window was intended to exclude consultations caused by COVID-19 symptoms for which testing was delayed by a few days, but to include those exposed to health care facilities in the time interval during which the infection was likely to have been acquired.  Our intention was to capture all cases that *could have been* nosocomial rather than the more restrictive definition  used in Scotland and England for "probable healthcare -associated infection" as illness onset (or first positive specimen date at least 8 days after admission  which aims to  identify cases that are unlikely to have been infected outside hospital [@public_health_england_covid-19_2020-1].

## Statistical methods
The analyses presented here focus  mainly on severe COVID-19 as this case definition does not depend upon the extent of testing for SARS-CoV-2, which varied markedly over the study period and is also likely to differ between those eligible for shielding and those ineligible.  Care home residents were excluded from case-control analyses as shielding advice intended for individuals in private households would not have been relevant to care home residents.

Cases and controls were grouped into three categories: no risk condition, moderate risk condition but ineligible for shielding, and eligible for shielding. Rate ratios for severe COVID-19 associated with eligibility for shielding are calculated with "no risk condition" as reference category. All rate ratios were estimated by conditional logistic regression.  As the  community controls were drawn by incidence density sampling and matched for age, sex, and general practice, this controls for these variables and for calendar time.

To plot the time course of the rate ratios associated with eligibility for shielding and with hospital exposure, these rate ratios were estimated over `r winsize`-day sliding windows of calendar time.  The fraction of severe cases attributable to hospital exposure was calculated in each 21-day time window from Miettinen's formula as \ensuremath{p_c \left( r - 1 \right) / r}.  where \ensuremath{p_c} is the frequency of exposure in cases in that time window and \ensuremath{r} is the rate ratio [@hsieh_effect_1988].  Results were similar with use of Levin's formula \ensuremath{p \left( r - 1 \right) / \left[ 1 + p \left( r - 1 \right) \right]}  where \ensuremath{p} is the frequency of exposure in controls.

Sliding windows of size  `r winsize.casedates`, `r winsize.hosp`, and `r winsize.fatality` days were used to plot the number of severe cases, the frequency of recent hospital exposure, and the rate ratios associated with hospital exposure and household composition.  In the plots of exposure frequency and rate ratio the data points from 1 June 2020 to 30 September 2020 are omitted as the numbers of cases and controls are too low for estimates of frequencies and rate ratios to be accurate.

# Results
A methodological concern was that unrecorded filtering of fatalities while the list of those eligible for shielding was being prepared could have led us to misclassify early fatal cases among those eligible for shielding as ineligible, leading to underestimation of the rate of fatal disease in the first few weeks of the epidemic in those eligible for shielding. As described in the Supplementary Material, by examining the time course of the fatal case proportion we were able to establish that such misclassification was unlikely in cases presenting after the second week of March 2020.

## Shielding eligibility and risk of COVID-19
Table \ref{tab:shieldgroupsbyage} shows the frequency of risk categories in those eligible for shielding.  To comply with statistical disclosure control rules, the `r shielded.full[shield.group=="Pregnant with heart disease", .N]` women who were identified as in the category "pregnant with heart disease" have been omitted from this table and all subsequent analyses; there were no severe cases in this group.

The largest risk category was those with severe respiratory disease.   Solid organ transplant recipients and those on immunosuppressants were younger than those in the other categories. Supplementary Table \ref{tab:letterdates} shows the proportions in each of these risk categories by date of advice letter: most of the solid organ transplant recipients, those with severe respiratory disease, and cancer patients were included in the first batch of letters on April 3rd 2020, but most of those in other risk categories were included in later batches.

Table \ref{tab:shieldcasegroup} shows  among the shielding group the number in each eligibility category by case status classified as not diagnosed, not severe or severe COVID-19.

Overall among `r nrow(shielded.full)` persons eligible  for shielding `r table.shield.casegr["All shielding categories", 2]` were diagnosed with COVID-19 but did not enter critical care or die within 28 days, `r table.shield.casegr["All shielding categories", 3]` entered critical care for COVID-19 but survived, and `r table.shield.casegr["All shielding categories", 4]` had a fatal outcome. Those on immunosuppressants had the lowest risks.

Figure \ref{fig:plotcasesbydate} compares the time series of severe cases in the three risk groups as a ribbon plot.  The daily number of presentations of severe cases in the total population reached a peak on `r dateformat(peakdate)` and fell to one quarter of the peak by `r dateformat(deaths.quartermax.date)`.   After each peak, the number of daily cases fell more slowly in those eligible for shielding than in those without risk conditions.

## Rate ratio for severe COVID-19  by risk condition group   
Of those eligible for shielding who were included in the case-control dataset, `r cc.severe[CASE==1 & shield.any==TRUE, .N, by=care.home][care.home=="Care/nursing home", N]` of the `r cc.severe[CASE==1 & shield.any==TRUE, .N]` severe cases and `r cc.severe[CASE==0 & shield.any==TRUE, .N, by=care.home][care.home=="Care/nursing home", N]` of the `r cc.severe[CASE==0 & shield.any==TRUE, .N]` matched controls were resident in care homes.  In all subsequent case-control analyses care home residents and their matched controls have been excluded, as policies for shielding residents of care homes are different from those relevant to shielding individuals living independently.

Table \ref{tab:regressionseverecasesnocare} shows the rate ratios for severe COVID-19 associated with shielding eligibility and with presence of a moderate risk condition, with those without risk conditions as reference category.  The univariate rate ratio for severe disease was
`r format.estcipv(table.listedgr3.severecases.nocare["Moderate risk condition", 3], table.listedgr3.severecases.nocare["Moderate risk condition", 4])` 
in those with moderate risk conditions and
`r format.estcipv(table.listedgr3.severecases.nocare["Eligible for shielding", 3], table.listedgr3.severecases.nocare["Eligible for shielding", 4])` 
in those eligible for shielding.  Among those eligible for shielding, solid organ transplant recipients were the group at highest risk, with a univariate rate ratio of  `r format.estcipv(table.shielded.severecases.nocare["Solid organ transplant", 3], table.shielded.severecases.nocare["Solid organ transplant", 4])` for severe COVID-19.     On multivariate adjustment rate ratios for all these groups remained high.

### Time course of the rate ratio associated with eligibility for shielding 
Figure \ref{fig:plotrateratio} (upper panel) shows the time course of the univariate rate ratios for severe COVID-19 n the  shielding and moderate risk categories, with those without risk conditions as reference category.   These rate ratios are computed over `r winsize`-day windows of presentation dates.  The rate ratio associated with eligibility for shielding increased from `r round(coeffs.long[date.midpoint==as.Date("2020-04-01") & riskgroup=="Eligible for shielding", rateratio], 2)` in the time window with mid-point 1 April to `r round(coeffs.long[date.midpoint==as.Date("2020-05-01") & riskgroup=="Eligible for shielding", rateratio], 2)` in the time window with mid-point 1 May.  Therefore there was no evidence that  the  shielding letters reduced the rate ratio associated with shielding eligibility.  A similar trend of rate ratio over time was found for those with moderate risk conditions.

## Associations of severe COVID-19 with transmission setting  and occupation -related factors among those eligible for shielding and in the overall population
Supplementary Table \ref{tab:shieldedonlyregressionseverecasesnocare} shows associations with risk factors among those eligible for shielding only, with severe respiratory disease (the largest category) as reference category.  The rate ratio increased with the number of adults in the household  but not with the number of children.  The strongest risk factor for severe COVID-19 among those eligible for shielding was recent hospital exposure, with a rate ratio of `r format.estcipv(table.shieldedonly.severecases.nocare["Recent hospital visit/stay", 5], table.shieldedonly.severecases.nocare["Recent hospital visit/stay", 6])` in the multivariable model.  Among those eligible for shielding who were cases of severe COVID-19, the proportion with recent hospital exposure was `r table.shieldedonly.severecases.nocare["Recent hospital visit/stay", 2]` had a recent hospital exposure. The fraction of severe cases attributable to recent hospital exposure can thus be calculated as `r round(100 * pc * (r -1) / r)`%.

Table \ref{tab:regressionseverecasesnocare}  shows the association with risk factors in the case control study overall.  The strongest risk factor (in terms of deviance explained) for severe COVID-19 in the general population was also recent hospital exposure: rate ratio `r format.estcipv(table.shielded.severecases.nocare["Recent hospital visit/stay", 3], table.shielded.severecases.nocare["Recent hospital visit/stay", 4])`.   As shown in Figure \ref{fig:plotexphosp} (middle panel)  the rate ratio associated with recent hospital exposure increased from `r round(coeffs.hosp.timewindow[date==as.Date("2020-04-01"), rateratio], 1)` in the time window with mid-point 1 April to `r round(coeffs.hosp.timewindow[date==as.Date("2020-05-01"), rateratio], 0)` in the time window with mid-point 1 May. Of the `r cc.severe[CASE==1 & hosp.recent==TRUE, .N]` cases who were classified as having recent hospital exposure, only `r cc.severe[CASE==1 & hosp.recent==TRUE & hosp.recent==TRUE & prob.hcai==TRUE, .N]` met the criteria for probable health-care associated infection.

Table \ref{tab:regressionseverecasesnocare} also show that the risk of severe COVID-19 increased with the number of adults in the household, but was inversely associated with the number of school-age children in the household in a  multivariable model. The rate ratio associated with two or more adults (with single-adult households as reference category), was   `r format.estcipv(table.listedgr3.severecases.nocare["adultsgt1", 5], table.listedgr3.severecases.nocare["adultsgt1", 6])` and the rate ratio associated with one or more school-age children was `r format.estcipv(table.listedgr3.severecases.nocare.anyschoolage["hh.schoolage.any", 5], table.listedgr3.severecases.nocare.anyschoolage["hh.schoolage.any", 6])`.   The other demographic factor associated with increased risk was socioeconomic deprivation: the rate ratio in the least deprived quintile compared with the most deprived quintile was `r format.estcipv(table.shielded.severecases.nocare["5 - least deprived", 5], table.shielded.severecases.nocare["5 - least deprived", 6])`.

As shown in Figure \ref{fig:plotrateratio} (lower panel), in a joint model with number of adults in household, number of children in household, and SIMD deprivation score the rate ratio for severe disease associated with adults remained in the range 1.5 to 2 throughout the epidemic, and the rate ratio per school-age child in household remained mostly in the range 0.7 to 1.

As reported previously, in comparison with other occupations patient-facing health-care workers were at higher risk of severe disease with a rate ratio of `r format.estcipv(table.shielded.severecases.nocare["Health care PF", 3], table.shielded.severecases.nocare["Health care PF", 4])`, and teachers were at lower risk with a rate ratio of `r format.estcipv(table.shielded.severecases.nocare["Teacher", 3], table.shielded.severecases.nocare["Teacher", 4])`.

As shown in Figure \ref{fig:plotexphosp}  (top panel) the frequency of recent hospital exposure in controls fell when restrictions were imposed in March-April 2020 and November-December 2020, but remained higher in those eligible for shielding than in those with moderate risk conditions or no risk conditions. It is the frequency of exposure in controls that is relevant to estimating background exposure levels, as the frequency of exposure in cases depends on the rate ratio.  The population attributable risk fraction of severe cases associated with recent hospital exposure reached a peak in the first wave of `r round(100 * coeffs.hosp.timewindow[date==as.Date("2020-05-01"), popattr.frac])`% at the beginning of May, declined rapidly over the next few weeks, and increased again to `r round(100 * coeffs.hosp.timewindow[date==as.Date("2020-12-01"), popattr.frac])`%  at the beginning of December Figure \ref{fig:plotexphosp} (bottom panel).  For comparison, over the period of study the population attributable risk fraction of severe cases associated with the narrow definition of "probable health-care associated infection" was only `r round(100 * hcai.popattrfrac, 1)`%.

# Discussion
## Summary and interpretation of findings
Key findings of this study were: (i) that the rate ratios for  severe COVID-19 associated with shielding conditions remained high through the epidemic; (ii) that recent hospital exposure and  the number of adults in the household were important  risk factors for severe COVID-19 among those eligible for shielding and also in the general population; (iii) that among those not resident in care homes the about 40% of severe cases among those not resident in care homes were attributable to hospital attendance even in the second wave.

In  comparison with those with no risk condition, the highest rate ratio for severe COVID-19 was the `r round(as.numeric(gsub(" .*", "", table.shielded.severecases.nocare["Solid organ transplant", 3])))`-fold rate ratio in solid organ transplant recipients. For other conditions deemed eligible for shielding, the rate ratios were between 4 and 7, compared to the rate ratios of about 3 associated with conditions designated as moderate risk, which include heart disease, diabetes, chronic kidney disease and disabling neurological conditions.  Other studies of outcome of COVID-19 in solid organ transplant recipients and other individuals using immunosuppressants have been based only on patients admitted to hospital [@avery_inpatient_2020; @andersen_association_2021]. Such studies cannot assess the risks to immunosuppressed individuals in the population.

If the advice provided and support offered in the shielding letters had been effective (beyond the messages already widely disseminated by this time and  general social distancing measures that were introduced  by then), we would have expected the rate ratio for severe disease associated with eligibility for shielding to fall within two weeks of the first letters being sent out.  Our results show however that while presentations of severe cases fell rapidly in the general population from the beginning of April 2020, the fall in presentations among those eligible for shielding advice was delayed so that the rate ratio for severe disease associated with eligibility for shielding rose during April 2020, although the daily deaths fell during April 2020 both in those eligible and those ineligible for shielding.

Recent hospital exposure was strongly associated with disease in all risk groups.  As recent hospital exposure was more common in those with moderate risk conditions and those eligible for shielding, especially after a lockdown had been imposed, this exposure accounted for a higher proportion of severe cases in these groups.  Thus a possible explanation for this rise in the rate ratio for severe disease  associated with eligibility for shielding during April 2020 and the subsequent rise beginning in October 2020 is that those in this group were not able to avoid exposure to hospitals during periods when hospitals were a source of infection.

The association of severe COVID-19 with number of adults in the household is consistent with classic studies of other viral infections showing that secondary cases in a household, where the infecting dose is likely to be high, are more severe than index cases [@aaby_malnutrition_1988].  The inverse association of severe COVID-19 with number of school-age children in the household and the lower risk in teachers compared with others of the same age and sex are consistent with a protective effect of past exposure to children  Lower risk in teachers has been reported previously [@mcallister_report_2020; @ludvigsson_open_2021].  These associations are consistent with evidence that past infection with other coronaviruses generates cross-reactive T-cell responses that may confer some resistance to SARS-CoV-2 [@braun_sars-cov-2-reactive_2020]. However no direct studies of the protective effect of these T-cell responses have been reported.  Although the rate ratio per child has been higher in the second wave of the epidemic than in the first wave, it has remained below 1 in almost all time windows.

## Comparison with previous literature
Other reports have evaluated the theoretical possibilities for shielding but without empirical data on its effectiveness.  A previous report [@jani_comparison_2020] was based on the population of Greater Glasgow and Clyde health board. As no age-standardized rate ratios were reported it is difficult to compare our results with theirs.  In the OPENSAFELY cohort, the rate ratio for fatal COVID-19 associated with living with children aged 0-11 years was 0.75 [@forbes_association_2020] after adjusting for covariates, but there was no association of non-fatal COVID-19 requiring critical care with living with children in this age group, and no dose-response relationship was reported.

## Strengths and limitations 
Strengths of our study are the population coverage, the range of outcomes assessed, and the comprehensive linkage to medical records and demographic risk factors.  A limitation is that we do not have primary care data other than encashed prescriptions. Furthermore, as most immunosuppressant drugs are prescribed through hospitals where linkage to prescribing records is not yet possible, the risks associated with  specific immunosuppressant drug classes could not be investigated.  We have no data on help with daily activities from non-resident carers as another possible source of exposure of clinically vulnerable individuals attempting to shield themselves.

The calculation of the fraction of cases attributable to an exposure provides an upper bound on the predicted effect of removing that exposure.  The definition of "recent hospital exposure" for this study is intended to capture all those whose infection could have been acquired in hospital (sensitivity of 1), unlike the official definition of "probable health-care associated infection" which is intended to identify those whose infection was unlikely to have been acquired outside hospital (high specificity).  This inclusive definition is appropriate as the calculation of the fraction of cases in the population that are attributable to the exposure is valid if the classification of exposure has sensitivity of 1, even if the specificity is less than 1 [@hsieh_effect_1988]. The association of severe COVID-19 with recent hospital admission is likely to be confounded by pre-existing risk conditions. However adjusting for risk conditions in a multivariable analysis reduced the rate ratio associated with recent hospital exposure only slightly (from 11.5 to 9.5) which suggests that this confounding effect is modest in comparison to the causal effect of exposure to hospital.  The large increase in the rate ratio associated with recent hospital exposure during periods when a high proportion of inpatients were hospital cases also supports causality.

## Policy implications 
The data presented here have important policy implications for shielding programmes,  for ongoing attempts to reduce COVID-19 transmission, and for vaccination policy.

Regarding shielding programmes, we could not find any evidence that the programme *per se* reduced COVID-19 rates  above and beyond the factors leading to the general reduction in rates in the background society.  It is  relevant to examine why advice to shield, combined with offers of support for delivery of food and medicines, failed to protect some individuals who were identified as clinically extremely vulnerable. As a lockdown on the general population had been imposed on 23 March 2020, and individuals who considered themselves to be at high risk would have been likely to reduce their contact level before then, there were limited possibilities for risk in this group to be reduced further by advice to shield and offers of help in letters sent from 3 April onwards.  We have identified two sources of exposure that are associated with severe disease and cannot easily be avoided by those advised to shield: hospital attendance as in-patient or out-patient, and sharing a household with other adults. We recommend that consideration should be given to special measures for to allow these  vulnerable individuals to reduce their exposure.  This could include alternative ways of delivering care without hospital attendance, and support for other household members to co-isolate with the vulnerable individual.

<!---Another important question is whether the correct risk groups were included in the shielded list. --->The designation of groups as clinically extremely vulnerable was necessarily  undertaken before any systematic studies of associations with severe disease could have been conducted.  Our study confirms that all these groups have markedly elevated risks of severe COVID-19, except for pregnant women with heart disease where the numbers at risk are too small for this risk to be reliably estimated, and therefore that assigning priority to these groups is appropriate for a vaccination programme that is aimed at reducing mortality and limiting the requirement for critical care.  <!---The inability to estimate the risk in pregnant women with heart disease in this study should not alter the   vaccination  priority currently given to pregnant women with severe acquired or congenital heart disease. --->
However we note that others in the population who do not have one of these conditions may nonetheless have similarly elevated risk from multiple factors, for example people with poorly-controlled diabetes and multiple complications.  To identify vulnerable individuals eligible for shielding or vaccination, the list of risk conditions could be supplanted by  formal risk scores.

Regarding  the importance of this study for policies to reduce overall transmissions we highlight here the different picture that emerges when associations of COVID-19 with any recent hospital exposure are calculated, rather than associations with probable hospital-acquired infection as formally defined.  Our analysis shows that focusing only on definite nosocomial inpatient-acquired infection may underestimate the impact that reducing infection risks across hospital sites could have.

Our results emphasise the importance of household contacts  in transmission in those with risk conditions and raise the question of the priority that should be given to vaccinating  household contacts of vulnerable individuals.   The Joint Committee on Vaccination and Immunization did not recommend this in a report published in December 2020, stating that "at this time there are no data on the size of the effect of COVID-19 vaccines on transmission." [@joint_commiteee_on_vaccination_and_immunisation_joint_2020].  The Committee however recommended vaccination of  health and social care workers who are "at risk of transmitting infection to multiple vulnerable persons or other staff in a health care environment",  implying some expectation that vaccination would reduce transmission.

For most of those designated as clinically vulnerable, vaccination is expected to provide protection against severe COVID-19.  For solid organ transplant recipients -- the group at highest risk -- it is uncertain whether the vaccine will evoke  an immune response sufficient to be effective and also uncertain whether  vaccines that use a live vector can be used safely [@chong_comprehensive_2017].  On the basis of our results, we recommend that household members of solid organ transplant recipients should be offered vaccination, without waiting for  more definitive evidence of the effect of vaccines on transmission. Even with vaccination of  household contacts, this group will require help with shielding until until passive immunisation is available or the risk of transmission has been reduced to a very low level by herd immunity.

## Information governance
This study was approved under COVID-19 Rapid Data Protection Impact Assessment​ (DPIA) 20210023 that allows Public Health Scotland staff to link the datasets.  Datasets were de-identified before analysis.

## Data availability
The component datasets used here are available via the Public Benefits Privacy Panel for Health at https://www.informationgovernance.scot.nhs.uk/pbpphsc/ for researchers who meet the criteria for access to confidential data. All source code used for derivation of variables, statistical analysis and generation of this manuscript is available on  https://github.com/pmckeigue/covid-scotland_public

## Funding 
No specific funding was received for this study.

## Transparency declaration
PM, as the manuscript’s guarantor, affirms: that the manuscript is an honest, accurate, and transparent account of the study being reported; that no important aspects of the study have been omitted; and that any discrepancies from the study as originally planned and registered have been explained. This manuscript has been generated directly from the source data by a reproducible research pipeline. All source code used for derivation of variables, statistical analysis and generation of this manuscript is freely available on GitHub.

# Registration
The original protocol for the REACT-SCOT case-control study was registered with the European Network of Centres for Pharmacoepidemiology and Pharmacovigilance (ENCEPP number EUPAS35558).

## Conflicts of interest
The authors declare no conflicts of interest.

## Acknowledgements
We thank all staff in critical care units who submitted data to the SICSAG database, the Scottish Morbidity Record Data Team, the staff of the National Register of Scotland, the Public Health Scotland Terminology Services, the HPS COVID-19 Laboratory & Testing cell and the NHS Scotland Diagnostic Virology Laboratories, and Nicola Rowan (HPS) for coordinating this collaboration.

## Public Health Scotland COVID-19 Health Protection Study Group
Allan McLeod​, Alice Whettlock April Went, Beth Findlay Chris Sullivan, Ciaran Harvey, David Henderson, Edward McArdle, Eisin McDonald, Emily Griffiths, Genna Drennan, Johanna Young, Kirstin Leslie, Leonardo I Green, Louise Nicol, Melissa Llano, Nick Christofidis, Paul Bett, Ross Cameron, Theresa Ryan, Victoria Ponce Hardy.  
Public Health Scotland, Meridian Court, 5 Cadogan Street, Glasgow G2 6QE

# References
<div id="refs"></div>

\newpage

# Figures
```{r plotcasesbydate, echo=FALSE, warning=FALSE, message=FALSE, fig.pos = "H", fig.asp=0.8, fig.cap="Ribbon plot of severe cases by vulnerability category and date of presentation, excluding care home residents"}

theme_set(theme_gray(base_size = 12))

p.byelig
```

```{r plotrateratio, echo=FALSE, warning=FALSE, message=FALSE, fig.asp=1.2, fig.cap="(a) Rate ratios for severe COVID-19 in those not resident in care homes, by risk group and sliding window of presentation dates.  Line thickness is proportional to precision of estimate. Arrows indicate dates that shielding advice letters were sent, with thickness proportional to number of letters sent on that date. (b) Rate ratio per adult and per child in household, in models with number of adults, number of children and SIMD deprivation score as covariates. Data points for dates from 1 June to 30 September 2020 are omitted because the numbers of cases and controls are small."}
gridExtra::grid.arrange(p.rateratio + theme(axis.title.x = element_blank()),
                        p.household.rateratio, nrow=2)
```

```{r plotexphosp, echo=FALSE, warning=FALSE, message=FALSE, fig.asp=1.2, fig.cap="(a) Recent hospital exposure by risk group and sliding windown of presentation dates (b) Rate ratio asociated with recent hospital exposure. (c) fraction of severe cases not resident in care homes attributable to recent hospital exposure. Data points for dates from 1 June to 30 September 2020 are omitted because the numbers of cases and controls are small."}
gridExtra::grid.arrange(p.hosp + theme(axis.title.x = element_blank()),
                        p.hosp.rateratio + theme(axis.title.x = element_blank()),
                        p.fracattr, nrow=3)

```

\newpage

# Tables
```{r shieldgroupsbyage, echo=FALSE, warning=FALSE, message=FALSE}
knitr::kable(table.shieldgroups.byage, 
             escape=FALSE, 
             booktabs=TRUE,
			 label="shieldgroupsbyage",
             row.names=TRUE,
             align=rep("r", 5),
             caption="Shielding eligibility cohort by eligibility category and age") %>%
    kableExtra::group_rows("Shielding eligibility category", 1, 6) %>%
        add_header_above(c(" "=1, "Age group"=4)) %>%
        add_footnote(label="Percentages are row percentages",
                     notation="none") %>%
       kable_styling(latex_options="HOLD_position")

```

```{r shieldcasegroup, echo=FALSE, warning=FALSE, message=FALSE}

knitr::kable(table.shield.casegr, 
             escape=FALSE, 
             booktabs=TRUE,
			 label="shieldcasegroup",
             row.names=TRUE,
             align=rep("r", 5),
             caption="Shielding eligibility cohort by case status. Severe cases are those that were fatal or entered critical care") %>%
    kableExtra::group_rows("Shielding eligibility category", 2, 7) %>%
    column_spec(column=1, width="4.5cm") %>%
    column_spec(column=2, width="2.5cm") %>%
    column_spec(column=3, width="2.2cm") %>%
    column_spec(column=4, width="2.2cm") %>%
    column_spec(column=5, width="2.2cm") %>%
    column_spec(column=6, width="2cm") %>%
        add_footnote(label="Percentages are row percentages",
                     notation="none") %>%
     kable_styling(latex_options=c("scale_down", "hold_position"))

```

\onehalfspacing

```{r regressionseverecasesnocare, echo=FALSE, warning=FALSE, message=FALSE}

knitr::kable(table.shielded.severecases.nocare,
             escape=FALSE,
             booktabs=TRUE,
             label="regressionseverecasesnocare",
             caption="Rate ratios for severe COVID-19 in those not resident in care homes. ", 
             col.names=c(clean.header(colnames(table.shielded.severecases.nocare)[1:2]),
                         "Rate ratio (95\\% CI)", "\\ensuremath{p}-value",
                         "Rate ratio (95\\% CI)", "\\ensuremath{p}-value")) %>%
    add_header_above(c(" "=3, "Univariate"=2, "Multivariable"=2)) %>%
    column_spec(column=1, width="3cm") %>%
    column_spec(column=2, width="2cm") %>%
    column_spec(column=3, width="2cm") %>%
    column_spec(column=4, width="2.5cm") %>%
    column_spec(column=5, width="2cm") %>%
    column_spec(column=6, width="2.5cm") %>%
    column_spec(column=7, width="2cm") %>%
    kableExtra::group_rows("Shielding eligibility category", 3, 8, bold=FALSE) %>%
    kableExtra::group_rows("Number of school age children in household", 10, 12, bold=FALSE) %>%
    kableExtra::group_rows("Number of adults in household", 13, 17, bold=FALSE) %>%
    kableExtra::group_rows("SIMD quintile", 18, 22, bold=FALSE) %>%
    kableExtra::group_rows("Occupation", 23, 26, bold=FALSE) %>%
    kable_styling(latex_options=c("scale_down", "HOLD_position")) %>%
    add_footnote(label=c("Percentages are column percentages for each variable",
                         "PF, patient-facing",
                         "Rate ratios are from conditional logistic regression models with cases and controls matched for age, sex and general care practice",
                         "Univariate rate ratios are for models with a single covariate",
                         "Multivariable rate ratios are for a model including all covariates"),
                 notation="none")

```

\newpage

\beginsupplement
# Supplementary Text
## Sensitivity analysis for misclassification of early fatal cases as ineligible for shielding
As those who were notified as having died were removed from the list of those eligible for shielding before letters were sent out, and no records of this filtering were available, we examined the fatal case proportion by presentation date.  The rationale for this was that as filtering the list of those eligible for shielding to remove those who had died would misallocate fatal cases to the ineligible for shielding category, we could estimate the date up to which this misallocation occurred by examining the fatal case proportion by presentation date.

Supplementary Figure \ref{fig:plotcasefatality} compares the time series of the fatal case proportion in those eligible and ineligible for shielding, by seven-day window of presentation date. The fatal case proportion in those eligible for shielding reached a peak in the first week of March and fell after this, suggesting that from mid-March onwards any misallocation of fatal cases to those ineligible for shielding was minimal. With a median 13 days from presentation to death, and further delay in death notifications, any filtering of notified deaths would have been unlikely to misallocate fatal cases whose presentation date was after the second week of March.

# Supplementary Figures

```{r plotcasefatality, echo=FALSE, warning=FALSE, message=FALSE, fig.asp=0.8, fig.cap="Proportion of diagnosed cases that were fatal in those ineligible and eligible for shielding, by sliding window of presentation date.  For each category, line thickness is proportional to number of cases"}

p.casefatality

```

\newpage

# Supplementary Tables

```{r letterdates, echo=FALSE, warning=FALSE, message=FALSE}

knitr::kable(table.date.letter,
             escape=FALSE, 
             booktabs=TRUE,
			 label="letterdates",
             row.names=TRUE,
             align=rep("r", 6), 
             caption="Dates of sending advice letters to those eligible for shielding") %>%
    column_spec(column=1, width="3cm") %>%
    column_spec(column=2, width="2.1cm") %>%
    column_spec(column=3, width="1.9cm") %>%
    column_spec(column=4, width="1.9cm") %>%
    column_spec(column=5, width="1.9cm") %>%
    column_spec(column=6, width="1.9cm") %>%
    add_footnote(label="Percentages are row percentages",
                 notation="none") %>%
    kable_styling(latex_options=c("scale_down", "hold_position"))

```

```{r shieldedonlyregressionseverecasesnocare, echo=FALSE, warning=FALSE, message=FALSE}

knitr::kable(table.shieldedonly.severecases.nocare,
             escape=FALSE,
             booktabs=TRUE,
             label="shieldedonlyregressionseverecasesnocare",
             caption="Rate ratios for severe COVID-19 associated with risk conditions (with severe respiratory disease as reference category) in those eligible for shielding and not resident in care homes", 
             col.names=c(clean.header(colnames(table.shieldedonly.severecases.nocare)[1:2]),
                         "Rate ratio (95\\% CI)", "\\ensuremath{p}-value",
                         "Rate ratio (95\\% CI)", "\\ensuremath{p}-value")) %>%
    add_header_above(c(" "=3, "Univariate"=2, "Multivariable"=2)) %>%
    column_spec(column=1, width="3.2cm") %>%
    column_spec(column=2, width="2cm") %>%
    column_spec(column=3, width="2cm") %>%
    column_spec(column=4, width="2.5cm") %>%
    column_spec(column=5, width="1.7cm") %>%
    column_spec(column=6, width="2.7cm") %>%
    column_spec(column=7, width="1.7cm") %>%
    kableExtra::group_rows("Shielding eligibility category", 1, 6, bold=FALSE) %>%
    kableExtra::group_rows("School-age children in household", 8, 10, bold=FALSE) %>%
    kableExtra::group_rows("Adults in household", 11, 13, bold=FALSE) %>%
    add_footnote(label=c("Percentages are column percentages for each variable",
                         "Rate ratios are from conditional logistic regression models matched for age, sex and general practice",
                         "Univariate rate ratios are for models with a single covariate",
                         "Multivariable rate ratios are for a model including all covariates"),
                 notation="none") %>%
    kable_styling(latex_options=c("scale_down", "hold_position"))

```