### 2021-08-29 under50norisk.Rmd

```{r freqsriskgroups, echo=FALSE, message=FALSE}
if(FALSE) {
knitr::kable(freqs.riskgroups, 
             escape=FALSE, 
             booktabs=TRUE,
			 label="freqsriskgroups",
             row.names=TRUE, 
			 align=c('r', 'r'), 
             caption="Numbers of controls and cases of severe COVID-19 from April to July 2021 by risk group and vaccination status") %>%
    add_header_above(c(" "=1, "Age 0-49 years"=2, "Age 50 years and over"=2)) %>%
    add_footnote(label=c("Presentation dates from 1 April 2021 to 14 July 2021.",
                         "Severe COVID-19 defined as entry to critical care or death within 28 days of presentation.", 
                         "Controls matched for age, sex, and primary care practice on date of presentation of case.", 
                         "Vaccination status is number of doses administered at least 14 days before presentation date"), 
                 notation="none") %>%
    kableExtra::group_rows("Vaccination status", 9, 11, bold=TRUE) %>% 
    column_spec(1, width="6cm") %>%
    column_spec(2, width="3cm") %>%
    column_spec(3, width="2.5cm") %>%
    column_spec(4, width="3cm") %>%
    column_spec(5, width="2.5cm") %>%
   kable_styling(latex_options=c("HOLD_position", "scale_down"))

##```

##```{r severebyagevax, echo=FALSE, message=FALSE}

knitr::kable(severe.byagevax, 
             escape=FALSE, 
             booktabs=TRUE,
			 label="severeagevax",
             row.names=TRUE, 
			 align=c('r', 'r'), 
             caption="Severe cases from April to July 2021 by age group and vaccination status") %>%
        add_header_above(c("Age group"=1, "Vaccine doses"=3)) %>%

    kable_styling(latex_options=c("HOLD_position"))

##```

##``{r casesnoriskchnums, echo=FALSE, message=FALSE}

knitr::kable(chnum.table, 
             escape=FALSE, 
             booktabs=TRUE,
			 label="casesnoriskchnums",
             row.names=FALSE, 
			 align=c('r', 'l', 'r', 'r'), 
             caption="Severe cases and controls aged under 50 without risk conditions presenting from April to July 2021: ICD-10 chapters represented at least once in discharge diagnoses  from 5 years to 25 days before presentation date") %>%
    column_spec(2, width="7cm") %>%
    add_footnote(label=c("Presentation dates from 1 April 2021 to 14 July 2021.",
                         "Table filtered to retain only ICD-10 chapters represented by at least 20 individuals.",
                         "The most frequent ICD subchapter in cases was F00 (alcohol-related disorders): 7 cases."),
                             notation="none") %>%
    kable_styling(latex_options=c("HOLD_position"))

##```

##```{r casesnoriskbnfchnums, echo=FALSE, message=FALSE}

knitr::kable(bnfchnum.table, 
             escape=FALSE, 
             booktabs=TRUE,
			 label="casesnoriskbnfchnums",
             row.names=FALSE, 
			 align=c('r', 'l', 'r', 'r'), 
             caption="Severe cases and controls aged under 50 without risk conditions presenting after March 2021: BNF chapters represented at least once in prescriptions dispensed during 240 to 15 days before presentation date") %>%
    column_spec(2, width="7cm") %>%
    add_footnote(label=c("Presentation dates from 1 April 2021 to 14 July 2021.",
                         "Table filtered to retain only BNF chapters represented by at least 20 individuals."),
                             notation="none") %>%
    kable_styling(latex_options=c("HOLD_position")) 
}
```