### 2021-08-29 vax2.Rmd

---
title: "Short report: efficacy of vaccination in those eligible for shielding in Scotland"
author:
  - name: Paul M McKeigue\textsuperscript{\getAff{Usher}}
    affiliation: HPS
  - name: David A McAllister\textsuperscript{\getAff{Glasgow}}  
    affiliation: HPS 
  - name: Jen Bishop
    affiliation: HPS
  - name: Sharon Hutchinson\textsuperscript{\getAff{GCU}} 
    affiliation: HPS
  - name: Chris Robertson\textsuperscript{\getAff{Strathclyde}}   
    affiliation: HPS
  - name: Nazir Lone
    affiliation: Usher
  - name: Jim McMenamin 
    affiliation: HPS
  - name: David Goldberg
    affiliation: HPS
  - name: Helen M Colhoun\textsuperscript{\getAff{IGMM}}
    corresponding: yes
    email: helen.colhoun@igmm.ed.ac.uk
    affiliation: HPS
address: 
- code: Usher
  address: Usher Institute, College of Medicine and Veterinary Medicine, University of Edinburgh, Teviot Place, Edinburgh EH8 9AG, Scotland. PM - Professor of Genetic Epidemiology and Statistical Genetics. NL - Clinical Senior Lecturer in Critical Care  
- code: HPS
  address: Public Health Scotland, Meridian Court, 5 Cadogan Street, Glasgow G2 6QE
- code: Glasgow
  address: Institute of Health and Wellbeing, University of Glasgow, 1 Lilybank Gardens, Glasgow G12 8RZ. DM - Senior Clinical Lecturer in Public Health 
- code: GCU
  address: School of Health and Life Sciences, Glasgow Caledonian University. SH - Professor of Epidemiology and Population Health 
- code: Strathclyde
  address: Department of Mathematics and Statistics, University of Strathclyde, 16 Richmond Street, Glasgow G1 1XQ. CR - Professor of Public Health Epidemiology
- code: IGMM
  address: Institute of Genetics and Molecular Medicine, College of Medicine and Veterinary Medicine, University of Edinburgh, Western General Hospital Campus, Crewe Road, Edinburgh EH4 2XUC, Scotland. HC - AXA Chair in Medical Informatics and Epidemiology

header-includes:
  \usepackage{newunicodechar}
  \let\origquote\quote
  \def\quote{\origquote\itshape}
  \usepackage{graphicx}
  \usepackage{geometry}
  \usepackage{longtable}
  \usepackage{booktabs}
  \usepackage{float}
  \floatplacement{figure}{H}
  \usepackage{array}
  \usepackage{threeparttable}
  \usepackage{longtable}
  \usepackage{lscape}
  \usepackage{pdfpages}
output: 
  rticles::nejm_article:
    includes:
      in_header: ./preamble.tex
    latex_engine: lualatex
    keep_tex: true
    fig_caption: yes
    md_extensions: -autolink_bare_uris
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
bibliography: ./covid.bib
csl: ./plos.csl
always_allow_html: true
urlcolor: blue
linenumbers: false
linkcolor: cyan
---

```{r vax, echo=FALSE, warning=FALSE, message=FALSE}

rollsum.datewin <- function(dt, k, datevar) {
    ## returns a table of rolling sums of width k centred on date
    ## dt is a dataset of totals (N) by date, which may have no rows for some dates in the
    ## range of interest
    ## add rows for missing dates by a left join of all.dates with dt
    all.dates <- data.table(date=seq(from=min(dt[[datevar]]),
                                     to=max(dt[[datevar]]),
                                     by=1))
    setkey(all.dates, date)
    # setkeyv(dt, datevar) can't set physical key here
    all.dates <- dt[all.dates]

    return(data.table(date=all.dates[[datevar]],
                      rsum=zoo::rollsum(all.dates[, N], k=k, fill=0, align="center")))
}

format.estcipv <- function(x.ci, x.pval) {
    x <- gsub(" \\(", " \\(95\\% CI ", as.character(x.ci))
    x <- gsub(", ", " to ", x)
    x <- gsub("\\)", paste0(", _p_=\\", x.pval, "\\)"), x)
    x <- gsub("times", "\\\\times", x)
    return(x)
}

format.estci <- function(x.ci) {
    x <- gsub(" \\(", " \\(95\\% CI ", as.character(x.ci))
    x <- gsub(", ", " to ", x)
    #x <- gsub("\\)", paste0(", _p_=\\", x.pval, "\\)"), x)
    #x <- gsub("times", "\\\\times", x)
    return(x)
}

options(knitr.kable.NA = '.')
options(knitr.table.format = "latex") # adds tab: prefix to labels
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)

## vaccination analysis

datadir <- "data/2021-03-16/"
load(file=paste0(datadir, "cc.all.RData"))

shield.bydate <- cc.all[CASE==0 & SPECIMENDATE > as.Date("2020-12-01") & shield.any==TRUE,
                           .N,
                           by=c("SPECIMENDATE")]
setkey(shield.bydate, SPECIMENDATE)
shield.bydate <- rollsum.datewin(shield.bydate, 7, "SPECIMENDATE")
setnames(shield.bydate, "rsum", "all")

if(FALSE) {
    shield.bydate <- cc.all[CASE==0 & SPECIMENDATE > as.Date("2020-12-01") & shield.any==TRUE,
                            .N,
                            by=c("SPECIMENDATE", "vax14.dose")]
    setkey(shield.bydate, SPECIMENDATE)
    vax0.bydate <- rollsum.datewin(shield.bydate[vax14.dose==0], 7, "SPECIMENDATE")
    vax1.bydate <- rollsum.datewin(shield.bydate[vax14.dose==1], 7, "SPECIMENDATE")
    vax2.bydate <- rollsum.datewin(shield.bydate[vax14.dose==2], 7, "SPECIMENDATE")
    setkey(vax0.bydate, date)
    setkey(vax1.bydate, date)
    setkey(vax2.bydate, date)
    setnames(vax0.bydate, "rsum", "rsum0")
    setnames(vax1.bydate, "rsum", "rsum1")
    setnames(vax2.bydate, "rsum", "rsum2")
    shield.bydate <- vax2.bydate[vax1.bydate]
    shield.bydate <- shield.bydate[vax0.bydate]
}

shieldvax.bydate <- cc.all[CASE==0 & SPECIMENDATE > as.Date("2020-12-01") &
                           shield.any==TRUE & vax14==1,
                           .N,
                           by=c("SPECIMENDATE")]
setkey(shieldvax.bydate, SPECIMENDATE)
shieldvax.bydate <- rollsum.datewin(shieldvax.bydate, 7, "SPECIMENDATE")
setnames(shieldvax.bydate, "rsum", "vax")
setkey(shield.bydate, date)
setkey(shieldvax.bydate, date)       
shield.bydate <- shieldvax.bydate[shield.bydate]
shield.bydate[, prop.vax := vax / all]

p.shieldvax <-
    ggplot(data=shield.bydate, aes(x=date, y=prop.vax)) +
    geom_line() +
    scale_y_continuous(breaks=c(0, 0.2, 0.4, 0.6, 0.8, 1), limits=c(0, 1),
                       expand=c(0, 0)) + 
    scale_x_date(breaks = seq.Date(from = as.Date("2020-12-15"),
                                   to =  as.Date("2021-03-02"), by = "week"),
                 labels=gsub("^0", "", 
                             format.Date(seq.Date(from = as.Date("2020-12-15"),
                                                  to =  as.Date("2021-03-02"),
                                                  by = "week"),
                                         "%d %b")
                             ),
                 limits=c(as.Date("2020-12-15"), as.Date("2021-03-02"))) +
        labs(x=paste0("Presentation date: mid-point of 7-day window"),
         y="Proportion vaccinated among controls eligible for shielding")

keep.vars <- c("SPECIMENDATE", "CASE", "discharge15to56", "prob.hcai", "care.home", "hh.over18", "listedgr3", "shield.any", "shield.group",
               listed.conditions, "shield.group", "vax14", "vax14.dose", "stratum")
cc.vax <- na.omit(cc.severe[SPECIMENDATE > as.Date("2020-12-01"), ..keep.vars])

keep.hosp.vars <- c("SPECIMENDATE", "CASE", "discharge15to56", "care.home", "shieldelig.group", 
                    "vax14", "stratum")
cc.hosp.vax <- na.omit(cc.all[(casegroup=="A" | casegroup=="B") &
                              SPECIMENDATE > as.Date("2020-12-01"), ..keep.hosp.vars])
rm(cc.all)

mh.strata <- cc.vax[care.home=="Independent" & shield.any==TRUE, 
                    .(SPECIMENDATE, stratum, CASE, vax14)]
mh.strata <- mh.strata[, c("a", "b", "c", "d") :=
                           list(CASE * as.integer(vax14),
                                (1 - CASE) * as.integer(vax14),
                                CASE * (1 - as.integer(vax14)),
                                (1 - CASE) * (1 - as.integer(vax14)))]
mh.strata <- mh.strata[, lapply(.SD, sum),
                       by=c("SPECIMENDATE", "stratum"), 
                       .SDcols=c("a", "b", "c", "d")]
mh.strata[, N := a + b + c + d]
mh.strata[, a.exp := (a + b) * (a + c) / N]
mh.strata[, c.exp := (a + c) * (c + d) / N]
mh.strata[, var.exp := (a + c) * (b + d) * (a + b) * (c + d) / (N^2 * (N - 1))]
mh.strata[is.nan(var.exp), var.exp := 0]
mh.strata[, week := paste(lubridate::year(SPECIMENDATE),
                          sprintf("%02d", lubridate::week(SPECIMENDATE)), sep=":")]

mh.dates <- mh.strata[, lapply(.SD, sum), by=week,
                      .SDcols=c("a", "a.exp", "c", "c.exp", "var.exp")]
setkey(mh.dates, week) 
mh.all <- data.table(week="All", mh.dates[, lapply(.SD, sum),
                                          .SDcols=c("a", "a.exp", "c", "c.exp", "var.exp")])
mh.dates <- rbind(mh.dates, mh.all)
colnames(mh.dates) <- c("Year:week", "Observed", "Expected", "Observed", "Expected", "Var(expected)")

table.dose <- summary(clogit(data=cc.vax, formula = CASE ~ care.home + listedgr3 / vax14.dose + strata(stratum)))$coefficients

table.riskgroups <- summary(clogit(data=cc.vax, formula = CASE ~ care.home + listedgr3 / vax14 + strata(stratum)))$coefficients
effect <- rownames(table.riskgroups)
table.riskgroups <- as.data.table(table.riskgroups)
table.riskgroups[, rateratio := or.ci(coef, `se(coef)`)]
table.riskgroups[, pvalue := format.pvalue(z, `Pr(>|z|)`)]
table.riskgroups <- table.riskgroups[, .(rateratio, pvalue)]
table.riskgroups <- data.table(effect=effect, # freqs.riskgroups,
                               table.riskgroups)
table.riskgroups[, effect := gsub("care.homeCare/nursing home", "Care home residence", effect)]
table.riskgroups[, effect := gsub("^listedgr3", "", effect)]
table.riskgroups[, effect := gsub("TRUE$", "", effect)]
table.riskgroups[, effect := gsub(":vax14$", "", effect)]
table.riskgroups[, effect := replace.names(effect)]

table.adjhosp <- summary(clogit(data=cc.vax, 
                               formula = CASE ~ care.home + listedgr3 / vax14 +
                                   prob.hcai + discharge15to56 + strata(stratum)))$coefficients

table.adjhosp <- as.data.table(table.adjhosp)
table.adjhosp[, rateratio := or.ci(coef, `se(coef)`)]
table.adjhosp[, pvalue := format.pvalue(z, `Pr(>|z|)`)]
table.adjhosp <- table.adjhosp[, .(rateratio, pvalue)]

table.shieldgroups <- summary(clogit(data=cc.vax, formula = CASE ~ care.home + shield.group / vax14 + strata(stratum)))$coefficients
effect <- rownames(table.shieldgroups)
table.shieldgroups <- as.data.table(table.shieldgroups)
table.shieldgroups[, rateratio := or.ci(coef, `se(coef)`)]
table.shieldgroups[, pvalue := format.pvalue(z, `Pr(>|z|)`)]
table.shieldgroups <- table.shieldgroups[, .(rateratio, pvalue)]
table.shieldgroups <- data.table(effect=effect, # freqs.shieldgroups,
                               table.shieldgroups)
table.shieldgroups[, effect := gsub("care.homeCare/nursing home", "Care home residence", effect)]
table.shieldgroups[, effect := gsub("^shield.group", "", effect)]
table.shieldgroups[, effect := gsub("TRUE$", "", effect)]
table.shieldgroups[, effect := gsub(":vax14$", "", effect)]
table.shieldgroups[, effect := replace.names(effect)]
table.shieldgroups <- as.data.frame(table.shieldgroups)

table.hosp <- summary(clogit(data=cc.hosp.vax, formula = CASE ~ care.home + shieldelig.group / vax14 + strata(stratum)))$coefficients
effect <- rownames(table.hosp)
table.hosp <- as.data.table(table.hosp)
table.hosp[, rateratio := or.ci(coef, `se(coef)`)]
table.hosp[, pvalue := format.pvalue(z, `Pr(>|z|)`)]
table.hosp <- table.hosp[, .(rateratio, pvalue)]
table.hosp <- data.table(effect=effect, # freqs.hosp,
                               table.hosp)
table.hosp[, effect := gsub("care.homeCare/nursing home", "Care home residence", effect)]
table.hosp[, effect := gsub("^shieldelig.group", "", effect)]
table.hosp[, effect := gsub("TRUE$", "", effect)]
table.hosp[, effect := gsub(":vax14$", "", effect)]
table.hosp[, effect := replace.names(effect)]
table.hosp <- as.data.frame(table.hosp)

freqs.riskgroups <- as.data.frame.matrix(cbind(
    univariate.tabulate(varnames=c("care.home", "listedgr3"), data=cc.vax,
                        drop.reflevel=FALSE),
    univariate.tabulate(varnames=c("care.home", "listedgr3"), data=cc.vax[vax14==1],
                        drop.reflevel=FALSE)))[-1, ]

oddsratio.table <- function(freqs.table) {
    freqs.table <- as.matrix(freqs.table)
    or <- numeric(nrow(freqs.table))
    for(i in 1:nrow(freqs.table)) {
        a <- as.numeric(gsub(" .+", "", freqs.table[i, ]))
        or[i] <- a[1] * a[4] / (a[2] * a[3])
    }
    return(round(or, 2))
}

freqs.shieldgroups <- as.data.frame.matrix(cbind(
   with(cc.vax[CASE==0], paste.rowpercent(table(shield.group, vax14))),
   with(cc.vax[CASE==1], paste.rowpercent(table(shield.group, vax14)))
   ))

freqs.hosp <- as.data.frame.matrix(cbind(
   with(cc.hosp.vax[CASE==0], paste.rowpercent(table(shieldelig.group, vax14))),
   with(cc.hosp.vax[CASE==1], paste.rowpercent(table(shieldelig.group, vax14)))
   ))

rownames(freqs.hosp) <- table.hosp$effect[8:14]

if(FALSE) {
cc.vax[`dm.typeOther/unknown type` == 1, `dm.typeType 2 diabetes` := 1]
cc.vax[, `dm.typeOther/unknown type` := NULL]

controls.vax <- cc.kept[CASE==0 & care.home=="Independent", .(AGE, COVID.age, vax_dose_1)]
controls.vax[, vaxnow := !is.na(vax_dose_1)]
controls.vax[, vaxnow := car::recode(vaxnow,
                                   recodes="FALSE='Not vaccinated';
                                            TRUE='At least one dose'",
                                   as.factor=TRUE,
                                   levels=c("Not vaccinated", "At least one dose"))]

theme_set(theme_gray(base_size = 16))
p.covidage <- ggplot(data=controls.vax, aes(x=AGE, y=COVID.age, color=vaxnow)) +
    geom_point(size=0.5, position="jitter") +
    scale_x_continuous(breaks=seq(20, 80, by=10), limits=c(15, 90)) + 
    scale_y_continuous(breaks=seq(20, 80, by=10), limits=c(15, 90)) +
    theme(legend.title = element_blank()) +
    theme(legend.position=c(0.15, 0.85)) +
    guides(color = guide_legend(override.aes = list(size = 3))) +
    labs(title=paste("COVID age versus calendar age by vaccination status on",
                        format.Date(max(vacc$vax_dose_1), '%d %b %Y')),
         caption="Controls matched to severe cases, not resident in care homes. Limits of COVID age are 16 and 90 years",
         x="Calendar age (years)",
         y="COVID age (years)")
p.covidage

png("COVIDage_vaxstatus.png")
p.covidage
dev.off()
rm(p.covidage)

vax.firstdoses <- vacc[, .N, by=c("weekdose1", "vax_type_1")]
ggplot(data=vax.firstdoses, aes(x=weekdose1, y=N, color=vax_type_1)) +
    geom_line()

}

```

# Key point summary

- This study examines the efficacy of vaccination in reducing the risk of severe COVID-19 among those designated as clinically extremely vulnerable and thus eligible for shielding in Scotland.
- This analysis is based on data from 1 December 1st 2020 to 16 March 2021 and therefore covers the initial dose of vaccine with very few second doses in the shielded having been administered as yet  in this period. 
- Efficacy of a single vaccination dose in protecting against severe COVID-19 was found to be as high or higher in those eligible for shielding, considered as a whole,  as in those without risk conditions.   
- Examination of  vaccine efficacy among those eligible for shielding broken down by the reason for shielding showed that for  solid organ transplant recipients the number of cases is still too small for the efficacy of vaccination to be estimated.  For all other groups there was a statistically significant reduction in hospitalisation for COVID-19 associated with  vaccination and that the magnitude of this reduction was similar to that in those who were not eligible for shielding. 
- Further data in larger populations are needed to evaluate vaccine efficacy in solid organ transplant recipients.
- Until such data are available it should not be assumed that vaccination is efficacious in this group.

# Background

In Scotland to date COVD-19 vaccination has been with either the ChAdOx1 nCoV-19 (Astra-Zeneca) or BNT162b2 mRNA (Pfizer-BioNtech) vaccines.  Both have been shown to be efficacious in clinical trials.  Real-world observational analysis in a primary care-based study in Scotland up to 15 Feb 15 2021 found maximal efficacy of 70% for ChAdOx1 nCoV-19 and 85% for BNT162b2 mRNA at 28 days post-vaccination [@vasileiou_effectiveness_2021].  However the efficacy of COVID-19 vaccination among those designated as clinically extremely vulnerable is still uncertain; an official [letter](https://www.gov.scot/publications/coronavirus-covid-19-shielding-letters/) sent on 23 March 2021 to those on this list stated that "we do not yet have evidence of exactly how effective [the vaccine] is for the people on the shielding list."   This is relevant to whether individuals in this category should be advised to remain in shielding as restrictions on the general population are lifted.  This study examines the efficacy of vaccination in reducing the risk of severe COVID-19 among those designated as clinically extremely vulnerable and thus eligible for shielding in Scotland.

# Methods 
This analysis is based on the REACT-SCOT matched case-control study, established at the beginning of the epidemic to explore risk factors for severe COVID-19 in the population of Scotland.  The design has been described in detail previously [@mckeigue_rapid_2020]. In brief, for every incident case of COVID-19 in the population ten controls matched to one-year age, sex and primary care practice were selected using the Community Health Index database.  COVID-19 cases are those with a positive nucleic acid test, or a hospital admission of death with COVID-19 codes. The dataset is refreshed regularly and is linked to the vaccination database and to the regularly updated dataset of all individuals deemed eligible for the shielding programme.  This analysis is based on all cases presenting between 1 December 1st 2020 to 16 March 2021.

We used this matched case-control design to take advantage of extensive data linkages already established. As the design uses incidence density sampling, it controls not only for the matching factors of age, sex and general practice but also for calendar time which would otherwise confound the association between vaccination and disease.

As previously, to minimise ascertainment bias we focused the analysis on  severe COVID-19,  defined as entry to critical care within 28 days of presentation or  fatal outcome (any death within 28 days of a positive test or any death for which COVID-19 was coded as underlying cause).   As described previously, cases and controls were grouped into one of three categories: no risk condition, risk condition designated previously by Public Health Scotland as moderate risk, eligible for shielding.  For further analyses, the shielding category was subdivided as described previously into six categories: solid organ transplant, specific cancers, severe respiratory conditions, rare conditions, on immunosuppressants, additional conditions.

Vaccination status in cases was coded as having had at least one dose of any vaccine at least 14 days before presentation date. This assigns cases (and their matched controls) who were vaccinated less than 14 days prior to the case date to the unexposed category.  The numbers of cases and controls were too few too allow further subdivision of time since vaccination.

The effect of vaccination in each of the clinical vulnerability categories was estimated in a conditional logistic regression model fitted to the full dataset, specifying effects $\beta_{R2}, \dots, \beta_{RJ}$ for the log rate ratio associated with risk categories 2 to $J$ ($\beta_{R1} = 0$ for the reference category $J=1$), and "nested" effects $\beta_{V1}, \dots, \beta_{VJ}$ for the log rate ratio associated with vaccination in each of the $J$ risk categories.  Because severe COVID-19 is strongly associated with care home residence and care home residents were a priority group for vaccination, the model includes the log rate ratio $\beta_H$ associated with care home residence. <!---For the $i$th individual in a matched set whose care home residence status is $H_i$, is in risk category $j$ and has vaccination status $V_i$, the linear predictor $\Theta_i$ is $\beta_H H_i + \beta_{Rj} + \beta_{Vj} V_{i}$.  The conditional likelihood for a matched set where individual $c$ is the case is $\exp{\left( \Theta_c \right) } / \sum_i{\exp{\left( \Theta_i \right) }}$. ---> The efficacy of vaccination in the $j$th risk category is 1 minus the rate ratio $\exp{\left(\beta_{Vj} \right)}$.

As recent hospital admissions or being in hospital before presentation date might confound the association of vaccination with severe COVID-19, the analysis was repeated with adjustment for any hospital discharge from 2 to 8 weeks before presentation date and for being an inpatient for at least 8 days before presentation date.   To allow inference on vaccine efficacy in less frequent risk groups, the analysis was also repeated with the case definition extended to include all cases that were hospitalized or fatal.  For this latter analysis the coding of moderate risk conditions was not available, so that those with no risk condition and those with moderate risk conditions are instead grouped together as "ineligible for shielding".

# Results
Figure \ref{fig:plotvax} shows the  proportion vaccinated at least 14 days before among those eligible for shielding in the control group, by presentation date.  This proportion in  controls  will be a close estimate of the  overall  vaccination rates in the total shielding population. From this we can infer that by 9 February 2021 about half the shielded population were vaccinated, rising to over 85% by early March.

```{r plotvax, echo=FALSE, warning=FALSE, message=FALSE, fig.pos = "H", fig.width=4, fig.asp=0.8, fig.cap="\\label{fig:plotvax}Proportion vaccinated at least 14 days before among those eligible for shielding in the control group, by presentation date"}

theme_set(theme_gray(base_size = 8))
p.shieldvax

```

Supplementary Table \ref{tab:freqsshieldgroups} shows  the vaccination status of  severe cases and  their controls  by risk stratum.  . Supplementary Table \ref{tab:freqshosp} shows the same breakdown for hospitalised cases and controls.   These   tables show that the numbers of cases in some shielding subgroups, especially solid organ transplant recipients, are too small for rate ratios to be estimated reliability. <!--delete the misleading crude odds ratios from these tables  alter the tables as discussed AND CHANGE THE TITLES ACCORDNGLY-->

The rate ratios for severe COVID-19 associated with vaccination were `r format.estci(table.riskgroups$rateratio[4])` in those with no risk condition, `r format.estci(table.riskgroups$rateratio[5])` in those with moderate risk conditions and `r format.estci(table.riskgroups$rateratio[6])` in those eligible for shielding.  With adjustment for having been in hospital for more than 7 days before presentation and for any hospital discharge from 15 to 56 days before presentation, the corresponding rate ratios were `r format.estci(table.adjhosp$rateratio[6])`, `r format.estci(table.adjhosp$rateratio[7])` and `r format.estci(table.adjhosp$rateratio[8])`.  
Table \ref{tab:shieldgroups} gives a breakdown by shielding group.  For the categories labelled as specific cancers, severe respiratory disease and additional conditions the numbers are large enough for the effect of vaccination on severe disease to be estimated reliably.  The point estimates for the rate ratio for all risk groups were  lower than for those without risk conditions i.e. vaccine efficacy was at least as high,  with the exception of  solid organ transplant recipients,  those  with rare diseases and  those on immunosuppressants.  For these groups the numbers of cases and controls were small (Supplementary Table \ref{tab:freqsshieldgroups}) and confidence intervals were wide.

```{r shieldgroups, echo=FALSE, warning=FALSE, message=FALSE}

knitr::kable(table.shieldgroups[9:16, ], 
             escape=FALSE, 
             booktabs=TRUE,
			 label="shieldgroups",
             row.names=FALSE,
             align=c("l", "r", "l"),
             caption="Rate ratios for severe COVID-19 associated with vaccination, within each risk group",
             col.names=c("Effect", 
                         "Rate ratio (95\\% CI)", "\\ensuremath{p}-value")) %>%
    add_footnote(label=c("Conditional logistic regression model matched on age, sex, general practice and presentation date",
                         "Vaccine effects nested within each level of risk group",
                         "Care home residence included as covariate", 
                         "Presentation dates from 1 December 2020 to 16 March 2021",
                         "Vaccine status coded as 1 if at least one dose at least 14 days before, 0 otherwise"), 
                 notation="none") %>%
    column_spec(column=1, width="6cm") %>%
    kable_styling(latex_options=c("HOLD_position"))

```

Table \ref{tab:shieldelig} shows the analysis extended to include all hospitalized cases, for which the numbers of cases are larger.   This shows significant efficacy in all groups other than solid organ transplant recipients but in this group the numbers of cases and controls were small even with the broader case definition (Supplementary Table \ref{tab:freqshosp}).

```{r hosp, echo=FALSE, warning=FALSE, message=FALSE}

knitr::kable(table.hosp[8:14, ], 
             escape=FALSE, 
             booktabs=TRUE,
			 label="shieldelig",
             row.names=FALSE,
             align=c("l", "r", "l"),
             caption="Rate ratios for hospitalized or fatal COVID-19 associated with vaccination within each risk group",
             col.names=c("Effect", "Rate ratio (95\\% CI)", 
                         "\\ensuremath{p}-value")) %>%
    add_footnote(label=c("Conditional logistic regression model matched on age, sex, general practice and presentation date",
                         "Care home residence included as covariate", 
                         "Presentation dates from 1 December 2020 to 16 March 2021",
                         "Vaccine status coded as 1 if at least one dose at least 14 days before, 0 otherwise"), 
                 notation="none")  %>%
    kable_styling(latex_options=c("HOLD_position"))

```

For more detail on the computation of the vaccine  efficacy in a matched case control design see  the statistical note in the supplement and the associated  Supplementary Table \ref{tab:mhdates} <!--this comes out as s table  4 but there is no S table 3 please rename--> .

# Conclusions

Efficacy of a single vaccination dose in protecting against severe COVID-19 is as high or higher in those eligible for shielding as in those without risk conditions.  For solid organ transplant recipients -- the group at highest risk among those eligible for shielding -- the number of cases is still too small for the efficacy of vaccination to be estimated.

The estimates of the rate ratio associated with vaccination or the vaccine efficacy here cannot be directly compared with the previous report for the total population of Scotland as they are summary estimates across all time periods from 14 days post-vaccination, rather than over 7-day intervals.  Nonetheless the estimate is close to that reported at 14 days in the  previous report.

Since we have previously shown that  the shielding group overall have approximately 6 fold higher risk of severe COVID-19 than those of the same age and sex without risk conditions [@mckeigue_relation_2021-1] they would need a vaccine efficacy of  82% (i.e. a rate ratio of 0.18)  with a single dose to achieve a lowering of risk to the level of unvaccinated persons  without risk conditions.   Thus whilst those in shielding  can be reassured that the vaccine is showing efficacy they should also be advised that they remain at increased risk  after a single dose (about twofold compared to those of the same age and sex without risk conditions given the vaccine-associated rate ratio of about 0.3). Thus they should remain vigilant  until they have received their second dose and background transmission rates are low.

Also a  larger dataset than the  Scottish population is needed to estimate efficacy in solid organ transplant recipients where we have not been able to demonstrate efficacy. It is important that  such an analysis be conducted so that such persons can be advised appropriately as to their ongoing risk level. We note that such data  would also be important to understand the need for development of potential alternatives  to vaccines such as synthetic antibodies that are now being evaluated in clinical trials [@eli_lilly_and_company_phase_2021]

The Vaccine Green Book recommends that the second dose of all  COVID-19 vaccines should be routinely scheduled between four and 12 weeks after the first dose [@public_health_england_covid-19_2021].  Although earlier rather than later second dosing within that interval might more quickly reduce the residual risk among those in shielding,   this is not necessarily the optimal strategy as the JCVI have noted that "With most vaccines an extended interval between the prime and booster doses leads to a better immune response to the booster dose." [@joint_committee_on_vaccination_and_immunisation_optimising_2021]

# References
<div id="refs"></div>

\beginsupplement
\onehalfspacing
\newpage

# Supplementary Material

```{r freqsshieldgroups, echo=FALSE, warning=FALSE, message=FALSE}

knitr::kable(freqs.shieldgroups, 
             escape=FALSE, 
             booktabs=TRUE,
			 label="freqsshieldgroups",
             row.names=TRUE,
				 col.names=c("Unvaccinated", "Vaccinated", "Unvaccinated", "Vaccinated"),
				 align=c("l", "r", "r", "r", "r"),
             caption="Numbers of controls and cases of severe COVID-19 by risk group and vaccination status") %>%
    add_header_above(c(" "=1, "Controls"=2, "Cases"=2)) %>% 
    add_footnote(label=c("Presentation dates from 1 December 2020 to 16 March 2021",
                         "Vaccine status coded as 1 if at least one dose at least 14 days before, 0 otherwise"), 
                 notation="none") %>%
    column_spec(2, width="2cm") %>%
    column_spec(3, width="2cm") %>%
    column_spec(4, width="2cm") %>%
    column_spec(5, width="2cm") %>%
    column_spec(6, width="1cm") %>%
    kable_styling(latex_options=c("hold_position"))

```

```{r freqshosp, echo=FALSE, warning=FALSE, message=FALSE}

knitr::kable(freqs.hosp, 
             escape=FALSE, 
             booktabs=TRUE,
			 label="freqshosp",
             row.names=TRUE,
  		 col.names=c("Unvaccinated", "Vaccinated", "Unvaccinated", "Vaccinated"),
             align=c("l", "r", "r", "r", "r"),
             caption="Numbers of controls and cases of hospitalized or fatal COVID-19 by risk group and vaccination status") %>%
    add_header_above(c(" "=1, "Controls"=2, "Cases"=2)) %>% 
    add_footnote(label=c("Presentation dates from 1 December 2020 to 16 March 2021",
                         "Vaccine status coded as 1 if at least one dose at least 14 days before, 0 otherwise"), 
                 notation="none")  %>%
    column_spec(2, width="2cm") %>%
    column_spec(3, width="2cm") %>%
    column_spec(4, width="2cm") %>%
    column_spec(5, width="2cm") %>%
    column_spec(6, width="1cm") %>%
    kable_styling(latex_options=c("hold_position"))

```

```{r mh.dates, echo=FALSE, warning=FALSE, message=FALSE}

knitr::kable(mh.dates[1:17, ], 
             escape=FALSE, 
             booktabs=TRUE,
			 label="mhdates",
             row.names=FALSE,
             align=c("l", "r", "r", "r", "r"),
             digits=1,
             caption="Vaccinated and unvaccinated severe cases among those eligible for shielding and not resident in care homes: observed and expected numbers by week") %>%
    add_header_above(c(" "=1, "Vaccinated cases"=2, "Unvaccinated cases"=2)) %>% 
    add_footnote(label=c("Expected numbers (under the null) are calculated in each stratum",
                         "Vaccine status coded as 1 if at least one dose at least 14 days before, 0 otherwise"), 
                 notation="none")  %>%
    kable_styling(latex_options=c("hold_position"))

```

\newpage

## Statistical note
Tables \ref{tab:freqsshieldgroups} and  \ref{tab:freqshosp}  show the the vaccination status of  distribution of cases and controls by shielding stratum  to illustrate that for solid organ transplant recipients the numbers of cases are too few to estimate vaccine efficacy reliably.   
We have used a matched case-control design, in which each incident case was matched to up to 10 controls of the same age (one-year band), sex and GP practice who were alive on the day of presentation of the case, forming a matched set.  Analyses of these matched sets are thus  inherently stratified by day.  Most matched sets have 1 case and up to 10 controls, but where two cases with the same age, sex and GP practice present on the same day the matched set would have 2 cases and up to 20 controls.

<!---
Each matched set can be represented as a 2 \ensuremath{\times} 2 table, in which the numbers in each cell are traditionally labelled $a, b, c, d$:

 | | Case | Control | All |
 | --- | --- | --- | --- |
 Vaccinated | $a$ | $b$ | $V_1$ |
 Unvaccinated | $c$ | $d$ | $V_0$ |
 All | $C_1$ | $C_0$ | $N$ |
--->
Put simply, the vaccine efficacy is estimated as the most likely value of the odds ratio associated with vaccination given the observed values in each matched set, conditional on the marginal totals (total cases, total controls, total vaccinated, total unvaccinated.  The log-likelihood of the odds ratio given the data is summed across all sets: matched sets in which all or none are vaccinated  do not contribute to the log-likelihood. So for example whilst it is obvious that sets that pertain to days before the vaccination programmes started cannot be informative it may be less obvious that sets pertaining to days when vaccination would be 100% complete are also uninformative. The calendar time period that is most informative is that when around 50% of the population   under study is vaccinated.

When estimating the efficacy within a risk group, only cases and controls in that risk group are retained in the matched sets, and only those only matched sets that contain both unvaccinated and vaccinated individuals contribute to the likelihood.

It follows from the above that crude tables of the vaccination status of cases and controls in each risk group cannot be used to estimate the efficacy.  This is especially true given that vaccination rates are rising sharply over the study period.

Table \ref{tab:mhdates} may give the reader a more intuitive  sense of how the information for computing vaccine efficacy in the shielded evolved over the study period.  It shows the weekly observed split of vaccinated versus unvaccinated cases among those in shielding versus the expected split conditional upon the marginal totals.  To keep the working simple, this table is restricted to those 778 cases eligible for shielding who  were not resident in care homes so there is no covariate adjustment.  The variance of the expected can be interpreted as the information about vaccine efficacy.