### 2021-08-29 vaxtimetrend.Rmd

---
title: "Time course of efficacy of vaccination against severe COVID-19 in Scotland"
author:
  - name: Paul M McKeigue\textsuperscript{\getAff{Usher}}
    affiliation: HPS
  - name: Helen M Colhoun\textsuperscript{\getAff{IGMM}}
#    corresponding: yes
#    email: helen.colhoun@igmm.ed.ac.uk
    affiliation: HPS
address: 
- code: Usher
  address: Usher Institute, College of Medicine and Veterinary Medicine, University of Edinburgh, Teviot Place, Edinburgh EH8 9AG, Scotland. PM - Professor of Genetic Epidemiology and Statistical Genetics. NL - Clinical Senior Lecturer in Critical Care  
- code: IGMM
  address: Institute of Genetics and Cancer, College of Medicine and Veterinary Medicine, University of Edinburgh, Western General Hospital Campus, Crewe Road, Edinburgh EH4 2XUC, Scotland. HC - AXA Chair in Medical Informatics and Epidemiology
- code: HPS
  address: Public Health Scotland, Meridian Court, 5 Cadogan Street, Glasgow G2 6QE

header-includes:
  \usepackage{newunicodechar}
  \usepackage[T1]{fontenc}
  \let\origquote\quote
  \def\quote{\origquote\itshape}
  \usepackage{graphicx}
  \usepackage{geometry}
  \usepackage{longtable}
  \usepackage{booktabs}
  \usepackage{float}
  \floatplacement{figure}{H}
  \usepackage{array}
  \usepackage{threeparttable}
  \usepackage{longtable}
output: 
  rticles::nejm_article:
    includes:
      in_header: ./preamble.tex
    latex_engine: lualatex
    keep_tex: true
    fig_caption: yes
    md_extensions: -autolink_bare_uris
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
bibliography: ./covid.bib
csl: ./f1000.csl # ./plos.csl
always_allow_html: true
urlcolor: blue
linenumbers: false
linkcolor: cyan
---

```{r vax, echo=FALSE, warning=FALSE, message=FALSE}
library(kableExtra)

options(knitr.kable.NA = '.')
options(knitr.table.format = "latex") # adds tab: prefix to labels
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)

```
\newpage

# Vaccine efficacy against Delta variant (or while Delta was the predominant variant)

| Study | Infection | Hospitalisation |  
| ----  | --------- | --------------- |  
| @nanduri_effectiveness_2021 | 53% |. | 
| @tenforde_sustained_2021 | . | 84% |  
| @rosenberg_new_2021 | 80% during July | 95%  |
| @lopez_bernal_effectiveness_2021 | Pfizer 88%, AZ 67% | . |  
| @puranik_comparison_2021 | Pfizer 42%, Moderna 76% | Pfizer 75%, Moderna 81% |  
| @pouwels_impact_2021 | Pfizer 80%, AZ 67% | . |

# References
<div id="refs"></div>

\newpage

# Figures

```{r bycalendartime, echo=FALSE, warning=FALSE, message=FALSE, fig.pos = "H", fig.width=6, fig.asp=0.8, fig.cap="\\label{fig:caseseverity} Time course of efficacy of 1 and 2 doses of each vaccine against severe disease: rate ratios in multivariable model admPoisson regression model for 90-day incidence of any case and severe case in the shielding cohort"}

theme_set(theme_gray(base_size = 10))

gridExtra::grid.arrange(p.rateratio, p.rateratio.hosp)

```